### 2020-01-29 article_4/newpaper.tex
report being revised following the comments and fixed ipynb error

\documentclass[]{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath,amssymb,amsthm} % For including math equations, theorems, symbols, etc
\usepackage{subfig}
\graphicspath{{Figures/}} 
\usepackage{hyperref}
\PassOptionsToPackage{dvipsnames}{xcolor}
\RequirePackage{xcolor} % [dvipsnames]
\definecolor{CTsemi}{gray}{0.55} % chapter numbers will be semi transparent .5 .55 .6 .0
\definecolor{CTcitation}{rgb}{0,0.5,0} % WebGreen
\definecolor{CTurl}{named}{Maroon} % Maroon
\definecolor{CTtitle}{named}{Maroon} % Maroon {cmyk}{0, 0.87, 0.68, 0.32}
\definecolor{CTlink}{named}{RoyalBlue} % RoyalBlue {cmyk}{1, 0.50, 0, 0}
\definecolor{halfgray}{gray}{0.55} % chapter numbers will be semi transparent .5 .55 .6 .0
\definecolor{webgreen}{rgb}{0,0.5,0}
\definecolor{webbrown}{rgb}{0.6,0,0}
\hypersetup{
	colorlinks=true,
	linkcolor=RoyalBlue,
	filecolor=magenta,      
	urlcolor=cyan,
	pdfkeywords={}
	breaklinks=true, bookmarks=true,bookmarksnumbered,
	urlcolor=webbrown, citecolor=webgreen, % Link colors
	pdftitle={}, % PDF title
	pdfauthor={\textcopyright}, % PDF Author
	pdfsubject={}, % PDF Subject
	pdfkeywords={}, % PDF Keywords
	pdfcreator={pdfLaTeX}, % PDF Creator
	pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}

%opening
\title{Bayesian Evidence Synthesis:opioid crisis}
\author{Hyeongcheol Park* \& Paul Gustafson* \& Micheal A Irvine*\textsuperscript{1}}

\begin{document}
%\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{#1}}}
\maketitle{}
\tableofcontents % Print the table of contents
\listoffigures % Print the list of figures
\listoftables % Print the list of tables

\begin{abstract}

\end{abstract}

\section{Introduction}
The opioid crisis is a major issue in North America  including Canada. There were 1,490 deaths and 15,598 paramedic- attended overdose events during 2017 alone. \cite{Irvine:modelling} (need to know about bib in latex, change statistics to 2018 later) The goal of this project is to apply Bayesian evidence synthesis to understand better the opoid crisis in Vancouver, Canada.  \\

[Need to give a backstory-why isn't this observable? The number of overdoses] 
[specify a geographic area]

All examples here were performed in Python 3.7 using the library pyMC (reference) and JAGS (reference). Training was performed using No U-Turn Sampling (NUTS) over two chains with 1000 iterations (is it sample size?). Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes. Data processing was carried out using the Pandas and SciPy library [reference]. Data visualization was performed using the libraries Seaborn and Matplotlib [ref]. Code for all examples in this study are provided.

[explain Bayesian statistics, MCMC and related topics]

\section{methods}

\normalsize
The number of overdoses is our ultimate interest of estimation.  Let $O_t$ the number of overdoses in a given month $t$. Suppose there was a survey  conducted to estimate the proportion of ambulance call $p_A$ among the overdoses. $p_A$ is assumed constant across time for simplicity. Let $n_{A}$ the sample size of the survey and $x_{A}$ to be the total number who confirmed they did call ambulance. It is assumed that $x_{A}$ follows a Binomial distribution:
[math font changing: suggest regular (not Greek letter) for observable]
\begin{equation}
\label{ambulance}
\left.\begin{aligned}
x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
\text{		ambulance call-outs model}
\end{equation}

The total overdoses need to be modeled. The simplest conceptual model is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$. \cite{Irvine:modelling} Denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$. It is assumed that the total overdose $O_t$ follows Poisson distribution where the population of the region of interest is $N$. 
[Give a graphical model representation? state that(?)]

\begin{equation}
\label{overdose}
\left.\begin{aligned}
z_{t} \sim N(\mu, \sigma^{2}) \\
\lambda_{t}^{OD} = \exp(z_{t})\\
O_{t} \sim Poi(\lambda_{t}^{OD}N) 
\end{aligned}\right\} 
\text{		overdose model} 
\end{equation}

[N, assumed known to investigator?]
Estimation of $O_t$ is not straightforward since none of the variables ($\mu$, $\sigma$, $N$) determining $O_t$  is known. Hence $O_t$ should be inferred from using $U_t$ and \(p_A\), where $p_A$ is the ambulance call out rate and \(U_t\) is the number of  ambulance-attended overdoses at a time point $t$. In general, the data of ambulance-attended overdoses \(U_t\) can be obtained. It is assumed that  \(U_t\) follows Binomial distribution: 
\begin{equation}
\label{over_amb}
\left.
U_t \sim Bin(O_t, p_A)
\right.
\end{equation}

Now $O_t$ can be estimated as $p_A$ can be infered by survey data and the data regarding $U_t$ is given. We suggest a simple model as a start where the model only combines Ambulance Call-outs Model (\ref{ambulance}) and Overdose Model (\ref{overdose}).

The next step is to run some simulations to figure out how different types of inputs lead some changes of output. To do so, a simple model is illustrated below.\\

\subsection{Simulation}

\normalsize 
The first simulation simplifies the assumptions of variables as much as possible; We assumed $N= 10000, n_{A}=1000$. The population size, N,  is unknown in reality but it is assumed known to investigator here for simplicity. The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; The total population size for a region, N, could vary over time or it can be staritified for a better realization of the real world. We vary the survey size  as $n_{A}=100$ or $n_{A}=10000$.  \\

\subsubsection{Likelihood}

\normalsize 
There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets are simulated as follows. 
The true value of $p_A$ was set $p_A=0.8$ for the survey data.
It is assumed that the data was collected for a year (t=1,2,3, ..., 12) and 
$x_t$ values were independentally generated from the Binomial distribution (\ref{ambulance}).  In terms of overdose data. It is assumed that the true values of parameters for overdose model were $\mu=\log0.05, \sigma=1$.
The vector of $O_t$ was generated following the overdose model (\ref{overdose}). The vector of $U_t$ was generated from the Binomial relation of the two variables ($\ref{over_amb}$). The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12). Then the 12 sueveys were combined into a single survey since $p_A$ is assumed fixed across time in this contexts. [Todo: generate 12 surveys and then combine them into one survey]

Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\

\subsubsection{Prior Distributions}
\normalsize Noninformative prior distributions are presumed as a start for simplicity.

\begin{equation}
\label{nonin_prior_amb}
p(p_A) \sim Beta(1,1)
\text{			noninformative prior of ambulance model}
\end{equation}

\begin{equation}
\label{noninprior_over}
\left.\begin{aligned}
\mu \sim U(-10,0)\\
\sigma \sim U(0,5)
\end{aligned}\right\} 
\text{			noninformative prior of overdose model}
\end{equation}

This leads the posterior distribution of variables of interest to heavily depend on the likelihood. Later, the noninformative priors will be changed and the impact of the changes over posteriors will be investigated. \\

\subsection{Early Result }

The result from the simple case scenario is illustrated below.

\subsubsection{Posterior Distribution}

Figure \ref{pst_ot} is the boxplot of posterior samples of $O_t$. It is shown that our posterior estimates of  $O_t$ are fairly accurate since (1) the boxplots contain actual values of $O_t$ within their interquartile range (IQR) and (2) the ranges of IQR and 95\% range seem narrow. Notice that the range of the boxplot from a higher $O_t$ values (t=4) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 4)\\

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{Figures/earlyresult1_ot.png}
	\caption{Boxplot of posterior samples of $O_t$ (2000 samples for each month) with actual data points of simulated $O_t$ values. The simulated values are shown as red dots.}
	\label{pst_ot}
\end{figure}

\newpage
\subsubsection{Posterior Predictive Check}

[Relationship between Ot and Ut makes sense?]
Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$. It is shown that the posterior predictive estimates of $U_t$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $O_t$. It is more obvious here that the range of the boxplots from  higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11)\\.

\begin{figure}[htb]
	\centering
	\subfloat[ppc of $U_t$]{\includegraphics[width=.50\columnwidth]{Figures/early_r_ppc1_ut.png}\label{ppc_ut}}
	\subfloat[ppc of $p_A$]{\includegraphics[width=.50\columnwidth]{early_r_ppc1_xt}\label{ppc_xt}}
	\caption[two early result box plots:]{Boxplots of posterior predictive samples of $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value. The simulated values are shown as a red dots.}

\end{figure}

Figure \ref{ppc_xt} is the boxplot of posterior predictive samples of $x_A$. It is shown that the posterior predictive estimates of $x_A$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $x_A$.\\

\subsection{Early Result: Contamination of $p_A$ } 
One of the attention of this research project is to investigate how robust the model is from contaminations of the data sets. The first inspection is to check an impact of a contamination of $p_A$; what would happen if the estimation of $p_A$ is biased? It is assumed that the survey data gives us a wrong estimate of $p_A$ such that it would be underestimated or overestimated. We then want to see how the biased estimation of $p_A$ affects the estimate of $O_t$, the total overdose.\\

Both of underestimation and overestimation were conducted for the analysis. In terms of underestimation, the simulated survey data ($n_A, x_A$) was generated with $p_A=0.6$ while the true value of $p_A$ is 0.8, and all the other assumptions hold the same. That is, $x_A$ is generated from $x_A \sim Bin(n_A, 0.6)$, while $U_t$ is generated from $U_t \sim Bin(O_t, 0.8)$ for every $t$. For overestimation, the simulated survey data was generated with $p_A=0.9$ while the true value of $p_A$ is 0.8, and all the other assumptions hold the same.  \\

\subsubsection{Posterior Distribution}

\begin{figure}[htb]
	\centering
	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-ot.png}\label{under_ot}}
	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-ot}\label{over_ot}}
	\caption[two early result box plots:]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $O_t$ values are shown as red dots.}
	\label{contam_ot}
\end{figure}

\normalsize 
Figure \ref{contam_ot} is the boxplots of posterior predictive samples where $p_A$ is contaminated from the survey data. It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots. This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); $O_t$ is generated by multiplying $U_t$ and the inverse of $p_A$ where $p_A$ is underestimated. This leads overestimated inverse of $p_A$ so that  $O_t$ becomes overestimated. Figure \ref{over_ot} shows the opposite case. Overestimation of $p_A$ leads underestimation of the inverse of $p_A$ which causes underestimation of $O_t$. From both figures it is seen that the bias increases as the estimated values and the actual values get large.\\

\subsubsection{Posterior Predictive Check}

\begin{figure}[htb]
	\centering
	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-ut.png}\label{under_ut}}
	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-ut}\label{over_ut}}
	\caption[two early result box plots:ut]{Boxplot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $U_t$ values are shown as red dots.}
	\label{contam_ut}
\end{figure}

Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that the none of the contaminations of $p_A$ leads an effect  $U_t$.\\

\begin{figure}[htb]
	\centering
	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-xa.png}\label{under_xa}}
	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-xa}\label{over_xa}}
	\caption[two early result box plots:xa]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $x_A$ values are shown as red dots.}
	\label{contam_xt}
\end{figure}

Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that the none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots. \\

However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900. This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); Only the ambulance model has an effect on the overdose model.

\bibliographystyle{unsrt}
\bibliography{sample.bib}
\end{document}

### 2020-01-29 article_4/structure.tex
report being revised following the comments and fixed ipynb error

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% Structure Specification File
%
% This file has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	REQUIRED PACKAGES
%----------------------------------------------------------------------------------------

\usepackage[
nochapters, % Turn off chapters since this is an article        
beramono, % Use the Bera Mono font for monospaced text (\texttt)
amsmath,% Use the Euler font for mathematics
pdfspacing, % Makes use of pdftex’ letter spacing capabilities via the microtype package
dottedtoc % Dotted lines leading to the page numbers in the table of contents
]{classicthesis} % The layout is based on the Classic Thesis style

\usepackage{arsclassica} % Modifies the Classic Thesis package

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs

\usepackage[utf8]{inputenc} % Required for including letters with accents

\usepackage{graphicx} % Required for including images
\graphicspath{{Figures/}} % Set the default folder for images

\usepackage{enumitem} % Required for manipulating the whitespace between and within lists

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{subfig} % Required for creating figures with multiple parts (subfigures)

\usepackage{amsmath,amssymb,amsthm} % For including math equations, theorems, symbols, etc

\usepackage{varioref} % More descriptive referencing

%----------------------------------------------------------------------------------------
%	THEOREM STYLES
%---------------------------------------------------------------------------------------

\theoremstyle{definition} % Define theorem styles here based on the definition style (used for definitions and examples)
\newtheorem{definition}{Definition}

 theorem styles here based on the plain style (used for theorems, lemmas, propositions)
\newtheorem{theorem}{Theorem}

\theoremstyle{remark} % Define theorem styles here based on the remark style (used for remarks and notes)

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%---------------------------------------------------------------------------------------

\hypersetup{
%draft, % Uncomment to remove all links (useful for printing in black and white)
colorlinks=true, breaklinks=true, bookmarks=true,bookmarksnumbered,
urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
pdftitle={}, % PDF title
pdfauthor={\textcopyright}, % PDF Author
pdfsubject={}, % PDF Subject
pdfkeywords={}, % PDF Keywords
pdfcreator={pdfLaTeX}, % PDF Creator
pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}

### 2020-01-29 article_4/newpaper.tex
dynamic figure saving fixed and done...

@@
 	\centering
-	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-ot.png}\label{under_ot}}
-	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-ot}\label{over_ot}}
+	\subfloat[ppc of $O_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-o_t.png}\label{under_ot}}
+	\subfloat[ppc of $O_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-o_t}\label{over_ot}}
 	\caption[two early result box plots:]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $O_t$ values are shown as red dots.}
@@
 	\centering
-	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-ut.png}\label{under_ut}}
-	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-ut}\label{over_ut}}
+	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-u_t.png}\label{under_ut}}
+	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-u_t}\label{over_ut}}
 	\caption[two early result box plots:ut]{Boxplot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
@@
 	\centering
-	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_under-xa.png}\label{under_xa}}
-	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_over-xa}\label{over_xa}}
+	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-x_a.png}\label{under_xa}}
+	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
 	\caption[two early result box plots:xa]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.

### 2020-01-29 article_4/newpaper.tex
minor changes of texts done again in latex

@@
 
-[Need to give a backstory-why isn't this observable? The number of overdoses] 
-[specify a geographic area]
+[Need to give a backstory-why isn't this observable? The number of overdoses] [specify a geographic area]
+\\
 
@@
 Code for all examples in this study are provided.
+\\
 
@@
 It is assumed that $x_{A}$ follows a Binomial distribution:
-[math font changing: suggest regular (not Greek letter) for observable]
+
 \begin{equation}
@@
 
-[N, assumed known to investigator?]
 Estimation of $O_t$ is not straightforward since none of the variables ($\mu$, $\sigma$, $N$) determining $O_t$  is known.
@@
 
-The next step is to run some simulations to figure out how different types of inputs lead some changes of output. To do so, a simple model is illustrated below.\\
+The next step is to run some simulations to figure out how different types of inputs lead some changes of output. To do so, the simple model is illustrated below.\\
 
@@
 The first simulation simplifies the assumptions of variables as much as possible; We assumed $N= 10000, n_{A}=1000$.
-The population size, N,  is unknown in reality but it is assumed known to investigator here for simplicity.
+Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity.
 The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; The total population size for a region, N, could vary over time or it can be staritified for a better realization of the real world.
@@
 $x_t$ values were independentally generated from the Binomial distribution (\ref{ambulance}).
-In terms of overdose data.
-It is assumed that the true values of parameters for overdose model were $\mu=\log0.05, \sigma=1$.
+In terms of overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
 The vector of $O_t$ was generated following the overdose model (\ref{overdose}).
@@
 The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).
-Then the 12 sueveys were combined into a single survey since $p_A$ is assumed fixed across time in this contexts.
-[Todo: generate 12 surveys and then combine them into one survey]
-
-Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\
+[Delete: Then the 12 sueveys were combined into a single survey since $p_A$ is assumed fixed across time in this contexts.] Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 It is shown that our posterior estimates of  $O_t$ are fairly accurate since (1) the boxplots contain actual values of $O_t$ within their interquartile range (IQR) and (2) the ranges of IQR and 95\% range seem narrow.
-Notice that the range of the boxplot from a higher $O_t$ values (t=4) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 4)\\
+Notice that the range of the boxplot from a higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1, 7, 11)\\
 
@@
 
-[Relationship between Ot and Ut makes sense?]
 Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$.
 It is shown that the posterior predictive estimates of $U_t$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $O_t$.
-It is more obvious here that the range of the boxplots from  higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11)\\.
+It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11).
+Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; For months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\
 
@@
 
-Figure \ref{ppc_xt} is the boxplot of posterior predictive samples of $x_A$.
-It is shown that the posterior predictive estimates of $x_A$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $x_A$.\\
+Figure \ref{ppc_xt} is the boxplots of posterior predictive samples of $x_A$.
+It is shown that the posterior predictive estimates of $x_A$ is tolerably accurate since (1) the boxplots contain actual values of $O_t$ within their lines connecting the maximum and the minimum and (2) the ranges of IQR seem modereately narrow.
+\\
 
@@
 \normalsize 
-Figure \ref{contam_ot} is the boxplots of posterior predictive samples where $p_A$ is contaminated from the survey data.
+Figure \ref{contam_ot} is the boxplots of posterior samples where $p_A$ is contaminated from the survey data.
 It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots.
-This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); $O_t$ is generated by multiplying $U_t$ and the inverse of $p_A$ where $p_A$ is underestimated.
-This leads overestimated inverse of $p_A$ so that  $O_t$ becomes overestimated.
+This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); $O_t$ is generated by multiplying $U_t$ and the inverse of $p_A$ where $p_A$ is underestimated; Hence $\frac{1}{p_A}$ is overestimated which leads overestimation of $O_t$.
+\begin{equation}
+\label{ot.how.made}
+\begin{aligned}
+O_t = U_t \frac{1}{p_A}
+\end{aligned}
+\end{equation}
 Figure \ref{over_ot} shows the opposite case.
@@
 Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that the none of the contaminations of $p_A$ leads an effect  $U_t$.\\
+It is seen that the none of the contaminations of $p_A$ leads an effect  $U_t$.
+[Explain why is this]\\
 
@@
 \end{figure}
-
 Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that the none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.
-\\
-
+It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.
 However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900.
-This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); Only the ambulance model has an effect on the overdose model.
+[I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); Only the ambulance model has an effect on the overdose model.]
 

### 2020-02-02 article_4/newpaper.tex
elaborating the paper_ adding figure

@@
 \section{Introduction}
-The opioid crisis is a major issue in North America  including Canada.
-There were 1,490 deaths and 15,598 paramedic- attended overdose events during 2017 alone.
-\cite{Irvine:modelling} (need to know about bib in latex, change statistics to 2018 later) The goal of this project is to apply Bayesian evidence synthesis to understand better the opoid crisis in Vancouver, Canada.
+The opioid crisis is a major issue in North America including British Columbia, Canada.
+British Columbia is in the midst of a drug overdose crisis due to illicit opioids and the circumstance is getting worse; According to the BC Coroners Service, there were more than 930 apparent illicit drug overdose deaths in BC from Jan.
+1 to Dec.
+31, 2016.
+This compares to 518 in 2015, an increase of 79.2\%.
+\cite{bccdc_opioid}  The goal of this project is to apply Bayesian evidence synthesis to understand better the opoid crisis in British Columbia, Canada.
 \\
 
-[Need to give a backstory-why isn't this observable? The number of overdoses] [specify a geographic area]
-\\
+One difficulty to cope with the opioid crisis is that the total number of overdoses is unknown.
+The reason is that the use of prescription opioids often leads to use illicit opioids and the illicit usage cannot be known.
+Many of those who become addicted to opioids do so after initially receiving a prescription.
+The highly addictive nature of these pain relievers makes it easy for the human brain to crave more.
+It is only after their prescription ends that many users realize they’ve become dependent on the effects of opioids  to function “normally.” At that point, they are either forced to get clean and endure the pain that comes with the withdrawal symptoms of opioids or look for another means of getting their high.
+This is often the time where people will turn to illicit drugs or other analogues.
+Because prescription opioids are so expensive, this is when many users turn to heroin.
+It is often cheaper, more potent, and easier to locate than what they were taking before.
+In fact, about 80\% of people using heroin started with a prescription to another opioid.
+After using heroin, however, 23\% of individuals develop opioid addiction.\cite{opioid_desc} [attempt: need to give a backstory-why isn't this observable?
+The number of overdoses]\\
+
+Since the total overdoses are unknown, the number needs to be estimated.
+Bayesian statistics can be a way to approach the problem and give us the good estimate of the number.
+Bayesian statistics is a theory in the field of statistics based on the Bayesian interpretation of probability where probability expresses a degree of belief in an event.
+\cite{wiki_bayes} Detailed explanation is provided at the following section.
+[attempt: explain Bayesian statistics]\\
 
@@
 
-[explain Bayesian statistics, MCMC and related topics]
-
 \section{methods}
@@
 \normalsize
+
 The number of overdoses is our ultimate interest of estimation.
-Let $O_t$ the number of overdoses in a given month $t$.
-Suppose there was a survey  conducted to estimate the proportion of ambulance call $p_A$ among the overdoses.
+To achive the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
+The posterior belief is coming from both parts of the information.
+Following equation express the idea in a mathmatical form:
+\begin{equation}
+\label{bayes_thm}
+\left.\begin{aligned}
+p(\theta| y ) = \frac{ p(\theta, y)   }{ p(y)} = \frac{ p(\theta) p(y| \theta)}{ p(y)} \propto p(\theta) p(y| \theta) \end{aligned}\right.
+\end{equation}
+where $\theta$ represents the target of estimation, the total overdoses and $y$ represents the collected data set providing relevant samples of the target variable.
+In Bayesian concepts, the variable of interest $\theta$ is a fixed number but the number is yielded from a certain distribution; That is, $\theta$ is considered to be a sample of a random variable.
+
+The problem, however, is that the data set $y$ is not available.
+Hence the technique is to approach the estimation in a indirect fashion.
+Figure \cite{simple_drawing} shows how we estimate the target variable.
+\\
+
+ Let $O_t$ the number of overdoses in a given month $t$ in British Columbia.
+[attempt: specify a geographic area] Suppose there was a survey  conducted to estimate the proportion of ambulance call $p_A$ among the overdoses.
 $p_A$ is assumed constant across time for simplicity.
@@
 
+[brief explaination of MCMC and related topics]
 \subsubsection{Likelihood}

### 2020-02-02 article_4/newpaper.tex
latex major revision done in the middle of Sunday

@@
 In Bayesian concepts, the variable of interest $\theta$ is a fixed number but the number is yielded from a certain distribution; That is, $\theta$ is considered to be a sample of a random variable.
-
-The problem, however, is that the data set $y$ is not available.
-Hence the technique is to approach the estimation in a indirect fashion.
-Figure \cite{simple_drawing} shows how we estimate the target variable.
+The problem, however, is that the data $y$ is not available.
+We can have our prior belief about the total overdoses but it is not possible to obtain any data.
+Hence it is needed to approach the estimation in a indirect fashion.
+Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
+[attempt: Give a graphical model representation?
+state that(?)]
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.5\linewidth]{Figures/simple_drawing}
+	\caption{Example of estimating the total overdoses indirectly given two data sets: ambulance attended overdoses and survey data. N is the population size}
+	\label{simple_draw}
+\end{figure}
 \\
@@
 x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
-\text{		ambulance call-outs model}
+\text{ambulance call-outs model}
 \end{equation}
@@
 It is assumed that the total overdose $O_t$ follows Poisson distribution where the population of the region of interest is $N$.
-[Give a graphical model representation? state that(?)]
 
@@
 \normalsize 
-The first simulation simplifies the assumptions of variables as much as possible; We assumed $N= 10000, n_{A}=1000$.
-Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity.
-The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; The total population size for a region, N, could vary over time or it can be staritified for a better realization of the real world.
-We vary the survey size  as $n_{A}=100$ or $n_{A}=10000$.
+The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.
+Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity as in figure \ref{simple_draw}.
+The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality.
+The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.
 \\
 
-[brief explaination of MCMC and related topics]
+\subsubsection{Markov chain Monte Carlo}
+The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
+The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived.
+[Further explanation is going to be added.] [attempt:brief explaination of MCMC and related topics]\\
+
 \subsubsection{Likelihood}
@@
 \subsubsection{Posterior Predictive Check}
+Posterior predictive checking is a model validation technique that we simulate some replicated data under the fitted model then compare the new data to the observed data.
+If the model fits, then replicated data generated under the model should look similar to observed data.
+To put it another way, the observed data should look plausible under the posterior predictive distribution.
+This is really a self-consistency check: an observed discrepancy can be due to model misfit or chance.
+\cite{bda_galman}
+[brief explannation of what is PPC]\\
 
@@
 It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11).
-Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; For months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\
+Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; for months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\
 
@@
 It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots.
-This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); $O_t$ is generated by multiplying $U_t$ and the inverse of $p_A$ where $p_A$ is underestimated; Hence $\frac{1}{p_A}$ is overestimated which leads overestimation of $O_t$.
+This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); given the likelihood of the ambulance attended overdoses, $U_t$, a value of $O_t$ is an unknown parameter of the binomial distribution, $Bin(O_t, p_A)$.
+Let $o_t$, $u_t$ are samples from $O_t$ and the likelihood value of $U_t$ respectively.
+Then $o_t$ is proportional to the quantity generated by multiplying $u_t$ and the inverse of $\hat{p}_{A}$ where $\hat{p}_{A}$ is underestimated estimator of $p_A$; hence $\frac{1}{\hat{p}_{A}}$ is overestimated which leads overestimation of $O_t$.
 \begin{equation}
@@
 \begin{aligned}
-O_t = U_t \frac{1}{p_A}
+o_t \propto u_t \frac{1}{\hat{p}_{A}}
 \end{aligned}
@@
 \subsubsection{Posterior Predictive Check}
+
+Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
+It is seen that none of the contaminations of $p_A$ leads an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
+
+The possible explanation is that the contaminations of  $p_A$ and $O_t$ (from $p_A$) cancel out the bias component as combined so that $U_t$ has no bias.
+Notice that $U_t$ is a likelihood (data set) of $O_t$ estimation.
+Hence it plays the role of $y$ in Bayes theorem (\ref{bayes_thm}) and the likelihood, $U_t$, is used to obtain the biased posterior samples of $O_t$ with the contaminated estimates of $p_A$.
+Here the pair of $(O_t, P_A)$ are both biased and the pair is fitted by the likelihood, $U_t$.
+To obtain the posterior predictive samples, MCMC algorithm help obtain posterior samples of biased $O_t$ and then the samples will be combined with biased $p_A$ such that the new samples of $\tilde{U_t}$ from $p(\tilde{U_t}|{U_t})$ is obtained.
+The result also matches with our intution.
+Posterior predictive checkes use the existing data points twice; firstly it uses the data to obtain the posterior samples of paramaters, and secondly the samples are used to produce posterior predictive samples.
+Given the fact that there could be an overfitting issue due to using the data twice, the result shown here seems reasonable.
+\\
 
@@
 
-Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that the none of the contaminations of $p_A$ leads an effect  $U_t$.
-[Explain why is this]\\
+Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
+It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.
+However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900.
+[I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.]
 
-\begin{figure}[htb]
+\begin{figure}[h]
 	\centering
@@
 \end{figure}
-Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.
-However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900.
-[I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); Only the ambulance model has an effect on the overdose model.]
 

### 2020-02-07 article_4/newpaper.tex
in the middle of the mist of pipeline building...

@@
 The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived.
-[Further explanation is going to be added.] [attempt:brief explaination of MCMC and related topics]\\
+\textbf{[Further explanation is going to be added.]} [attempt:brief explaination of MCMC and related topics]\\
 
@@
 \cite{bda_galman}
-[brief explannation of what is PPC]\\
+[attempted: brief explannation of what is PPC]\\
 
@@
 However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900.
-[I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.]
+[????
+I think this should go to the previous of the previous paragraph] This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.
 

### 2020-02-07 Rcodes/JAGS_practice.Rmd
mist of pipeline

---
title: "R Notebook"
output: html_notebook
---
First, we bring the two data sets (survey data and ambulance attended overdoses data) into here. It is done by manual labor for now but it should be automatical when the pipeline with pymc3 code is built (probably with reticulate package).
```{r}
df <- data.frame(
  o_t= c(2475,  262,  318,  149, 1151,   39, 2997,  242,  702,  377, 2088,76), 
  u_t=c(1969,  217,  253,  119,  934,   34, 2392,  196,  569,  308, 1655,55), 
  x_a= c(799, 798, 795, 816, 805, 794, 793, 780, 773, 779, 788, 813)
)
n_a=1000
N= 10000
```

runjags package

Todo:

1. Import relavent functions from the pymc3 Jupyter notebook.
2. Obtain the same data sets.

To do 1, I need to save the functions into a module or a library.
```{r Package preparing}
# package connecting R and Python
install.packages("reticulate", dependencies = TRUE)
# JAGS
install.packages('rjags')
library(reticulate)
library(rjags)
```
```{r}
py_install("pandas")
py_run_file("../report_revised_2.py")
```

```{r}
install.packages('RSQLite')
library(RSQLite)
```
```{r}
conn <- dbConnect(RSQLite::SQLite(), "notebook_env.db")
alltables = dbListTables(conn)
```

```{r}
install.packages('ProjectTemplate')
library('ProjectTemplate')
```

```{r}
db.reader('notebook_env.db', '../notebook_env.db', 'data')
```

### 2020-02-07 report_revised_2.py

@@ -1,0 +1,481 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+# # Basic Model: Ambulance + Overdose
+# 
+# 
+# 
+# ## Ambulance Call-outs Model
+# 
+# $n_{A}$: sample size   
+# $x_{A}$: the total number who confirmed they did call an ambulance  
+# $p_{A}$: probability of a person call an ambulance   
+# 
+# $$x_{A} \sim Bin(n_{A},p_{A})$$
+# 
+# **We assume $n_{A}=1000, p_A = 0.8$.**  
+# **Suppose the prior of $p_A$ is noninformative.** 
+# $$p(p_A) \sim Beta(1,1)$$
+# 
+# ## Overdose Model
+# Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**
+# Also, we have priors.
+# 
+# 
+# $$z_{t} \sim N(\mu, \sigma^{2})$$  
+# $$\lambda_{t}^{OD} = \exp(z_{t})$$  
+# $$O_{t} \sim Poi(\lambda_{t}^{OD}N)$$  
+# $$U_t \sim Bin(O_t, p_A)$$  
+# For simplicity we set N =10000 for now. We need to generate reasonable $U_t$ values first. Note that $U_t$ comes from $\mu, \sigma$ following all the way through the overdose model.  
+# 
+# 
+# 
+# $\mu=\log0.05, \sigma=1, N=10000$.   
+# We suppose survey data exists: ($n_A, x_A$) known.  
+# 
+# **We set for our prior parameters:** $$\mu \sim U(-10,0)$$  $$\sigma \sim U(0,5)$$
+
+# In[1]:
+
+
+import pandas as pd
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import pymc3 as pm
+import theano.tensor as tt
+import seaborn as sns
+
+
+# In[2]:
+
+
+import platform
+platform.python_version()
+
+
+# Below is the function that generate data from simulation.
+
+# In[3]:
+
+
+def sim_ambulance_overdoses(random=1, mu_z=np.log(0.05), sigma_z=1.,
+                            N=10000, p_a=0.8, n_a=1000, n_T=12):
+    '''
+    Simulate overdoses data, ambulance-call outs survey data and related variables 
+    -----------
+    mu_z- float 
+        log rate for lambda
+    
+    sigma_z -float
+        variance of log lambda
+    
+    N - int
+        Total size of population
[…]

### 2020-02-08 Rcodes/JAGS_practice.Rmd
jags lagged

@@
 ---
+# Following an online tutorial
+
+Context:Normal model with unknown mean and variance.
+
+```{r}
+install.packages('runjags')
+library(runjags)
+modelString = "
+model{
+for (i in 1:N) {
+y[i] ~ dnorm(mu, phi)
+} 
+mu ~ dnorm(mu_0, phi_0)
+phi ~ dgamma(alpha, beta)
+}
+"
+
+```
+- note that they take precison for Normal distribution
+-use 'd' to describe density: 'r' never used.
+
+```{r}
+y <- CEdata$LogTotalExpLastQ
+N <- length(y)
+the_data <- list(
+  'y'=y,
+  'N'=N,
+  'mu_0'=5,
+  'phi_0'=1/1^2,
+  'alpha'=1,
+  'beta'=1
+)
+
+```
+
+# Opioid Overdose Modeling 
 First, we bring the two data sets (survey data and ambulance attended overdoses data) into here.

### 2020-02-09 Rcodes/JAGS_practice.Rmd
jags lagged: errors to be questioned

@@
 ---
-# Following an online tutorial
+# Following an online tutorial (no data...)
 
@@
 
+summary: 
+in the model, as a string, we put the likelihood and the parameters of interest with their prior distribution.
 ```{r}
 install.packages('runjags')
+install.packages('coda')
+```
+
+```{r}
 library(runjags)
@@
 
+From here, the list named as the_data, we plug in the values of the parameters for the prior including the number of observation for likelihood (N).
+
+Posterior result is obtained by run.jags function where we plug in our data (the_data) which we set the values of the parameters for prior, we state our interest parameters as monitor.
+I assume modelString goes into the first argument here.
 ```{r}
@@
   'alpha'=1,
-  'beta'=1
+  'beta'=1) # this we where the pre-set priors given to the model
+  posterior <- run.jags(
+    'norm_model.jags',
+    data = the_data,
+    monitor = c('mu','phi'),
+    n.chains=1,
+    adapt=1000,
+    burnin=2000,
+    sample=5000,
+    thin=1
   )
 
+summary(posterior)
+
 ```
+
+# Another JAGS Tutorial 
+[from here](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html)
+
+```{r}
+setwd("../../meta-analysis-Bayesian/")
+data =read.csv('data1.csv')
+head(data)
+```
+
+```{r Defining values}
+library(coda)
+N <- length(data$y)
+
+data <- list('N'=N, 
+             'y'=data$y,
+             'V'=data$V)
+
+inits <- list(d=0.0)
+
+```
+
+```{r Enter the Model}
+cat("model{
+for (i in 1:N) {
+P[i] <- 1/V[i]
+y[i]~dnorm(d,P[i])
+}
+# Define the priors
+d ~ dnorm(0,0.00001)
+
+# Transfomr the ln(OR) to OR
+OR <- exp(d)
+}
+
+    ", file='model_test.txt')
+```
+
+```{r}
+library(rjags)
+jags.m <- jags.model(file='model_test.txt',data=data, inits=inits, n.chains = 1, n.adapt = 500,)
+```
+```{r}
+params <- c('d','OR')
+samples <- coda.samples(jags.m, params, n.iter=10000)
+```
+
+```{r}
+summary(samples)
+```
+
+```{r}
+summary(window(samples, start=5001))
+```
+
+```{r}
+plot(samples)
+```
+
+_______________________________________________________________
 

### 2020-02-09 Rcodes/JAGS_overdoses.Rmd
questions : JAGS

---
title: "JAGS_overdoses"
author: "Tom Park"
date: '2020-02-08'
output: html_document
---
# Basic Model: Ambulance + Overdose
## Ambulance Call-outs Model
$n_{A}$: sample size   
$x_{A}$: the total number who confirmed they did call an ambulance  
$p_{A}$: probability of a person call an ambulance   
$$x_{A} \sim Bin(n_{A},p_{A})$$
**We assume $n_{A}=1000, p_A = 0.8$.**  
**Suppose the prior of $p_A$ is noninformative.** 
$$p(p_A) \sim Beta(1,1)$$
## Overdose Model
Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**

Also, we have priors.
$$z_{t} \sim N(\mu, \sigma^{2})$$  
$$\lambda_{t}^{OD} = \exp(z_{t})$$  
$$O_{t} \sim Poi(\lambda_{t}^{OD}N)$$  
$$U_t \sim Bin(O_t, p_A)$$  
For simplicity we set N =10000 for now. We need to generate reasonable $U_t$ values first. Note that $U_t$ comes from $\mu, \sigma$ following all the way through the overdose model.  
$\mu=\log0.05, \sigma=1, N=10000$.   
We suppose survey data exists: ($n_A, x_A$) known.

**We set for our prior parameters:** $$\mu \sim U(-10,0)$$  $$\sigma \sim U(0,5)$$

```{r message=FALSE, warning=FALSE}
# install packages
install.packages('runjags')
install.packages('rjags')
install.packages('coda')
library('runjags')
library('rjags')
library('coda')
```
The data is the same data from pymc3 with Python.
```{r}
df <- read.csv('./basic_data.csv',check.names=FALSE)
head(df)
```
Now we set the model which defines the relations of overdose model and ambulance call model.
```{r Enter the models}
cat("model{
## define the priors
p_a <- dbeta(alpha, beta)
mu <- dunif(mu_a, mu_b)
sigma <- dunif(sigma_a, sigma_b)

## the latent variables
z <- dnorm(mu, 1/sigma^2)
lmd <-  exp(z)

for (i in 1:N) {
x_a[i] <- dbin(p_a, N) # each survey result for month
o_t[i] <- dpois(lmd) # total overdoses per month
u_t[i] <-dbin(p_a, o_t[i]) # ambulanced overdoses per month
}
}

    ", file='basic_model.txt')
```

```{r Defining Values}
N <-  length(df$o_t) # we have 12 samples
u_t <- df$u_t
x_a <- df$x_a
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' = 1,

            # priors for overdose model 
            'mu_a'=-10,
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=u_t, 
            'x_a'=x_a, # We do not include o_t since unknown
            "N"=N # total months
            )

```

I referred two tutorials online. 
[first one](https://www.youtube.com/watch?v=F9jxvhXc_1w) used run.jags for getting posterior distribution and [the second one](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html) used jags.model function.

I tried both and they didn't work.

This is using jags.model function from rjags package.
```{r Set Up the JAGS Model:jags.model}
jags.simple <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains=2,
                          n.adapt=1000)
```

# Q1.

Here I am confused since dunif is what I found correct [here](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf).

# Q2. 
When I tried the other package and function, I got an error that object "NA" not found as below.

This is using jags.model function from runjags package.
```{r another try:run.jags}
jags.simple <- run.jags(file='basic_model.txt',
                          data=dat,
                          monitor =c('u_t'),
                          n.chains=2,
                          n.adapt=1000)
```
```

### 2020-02-16 Rcodes/JAGS_overdoses.Rmd
JAGS DEBUG DONE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@
 ---
-title: "JAGS_overdoses"
+title: "JAGS Model"
 author: "Tom Park"
@@
 ---
-# Basic Model: Ambulance + Overdose
-## Ambulance Call-outs Model
+## Basic Model: Ambulance + Overdose
+### Ambulance Call-outs Model
 $n_{A}$: sample size   
@@
 $$p(p_A) \sim Beta(1,1)$$
-## Overdose Model
+### Overdose Model
+
 Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**
@@
 # install packages
-install.packages('runjags')
-install.packages('rjags')
-install.packages('coda')
+if (!require(runjags)) install.packages("runjags")
+if (!require(rjags)) install.packages("rjags")
+if (!require(coda)) install.packages("coda")
+
 library('runjags')
@@
 The data is the same data from pymc3 with Python.  
+Todo: build a pipeline to connect the python (pymc3) and R (JAGS)
+
 ```{r}
-df <- read.csv('./basic_data.csv',check.names=FALSE)
+df <- read.csv('./basic_data.csv')
+df$X <- NULL 
 head(df)
@@
 Now we set the model which defines the relations of overdose model and ambulance call model.
+
+The model defined as follows.
 ```{r Enter the models}
@@
 ## define the priors
-p_a <- dbeta(alpha, beta)
-mu <- dunif(mu_a, mu_b)
-sigma <- dunif(sigma_a, sigma_b)
+p_a ~ dbeta(alpha, beta)
+mu ~ dunif(mu_a, mu_b)
+sigma ~ dunif(sigma_a, sigma_b)
 
 ## the latent variables
-z <- dnorm(mu, 1/sigma^2)
-lmd <-  exp(z)
+z ~ dnorm(mu, 1/(sigma^2))
+lambda <- exp(z)
 
-for (i in 1:N) {
-x_a[i] <- dbin(p_a, N) # each survey result for month
-o_t[i] <- dpois(lmd) # total overdoses per month
-u_t[i] <-dbin(p_a, o_t[i]) # ambulanced overdoses per month
+for (i in 1:n) {
+  ## ambulance model 
+  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
 }
+for (i in 1:n) {
+  ## overdose model
+  o_t[i] ~ dpois(lambda*N) # total overdoses per month
+}
+for (i in 1:n) {
+  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
 }
 
-    ", file='basic_model.txt')
+}", file='basic_model.txt')
 ```
 
+Pre-set variables.
 ```{r Defining Values}
-N <-  length(df$o_t) # we have 12 samples
+n <-  length(df$o_t) # we have 12 samples
+n_a <- 1000
+N <- 10000
 u_t <- df$u_t
 x_a <- df$x_a
+```
+
+Define the list providing the values of the variables and the parameters for the priors of the model. 
+```{r from variable name to \'variable_name\'}
 dat <- list(
@@
             # priors for overdose model 
-            'mu_a'=-10,
+            'mu_a'=(-10),
             'mu_b'=0,
@@
             'u_t'=u_t, 
-            'x_a'=x_a, # We do not include o_t since unknown
-            "N"=N # total months
+            'x_a'=x_a,  
+            'N'=N, # the population
+            'n'= n, # total months
+            'n_a'=n_a
+
             )
-
+```
+**Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
+``` {r run the first model without error }
+jags.simple <- jags.model(file='basic_model.txt',
+                          data=dat)
 ```
 
-I referred two tutorials online. 
-[first one](https://www.youtube.com/watch?v=F9jxvhXc_1w) used run.jags for getting posterior distribution and [the second one](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html) used jags.model function.
-
-I tried both and they didn't work.
-
-This is using jags.model function from rjags package.
-```{r Set Up the JAGS Model:jags.model}
-jags.simple <- jags.model(file='basic_model.txt',
-                          data=dat,
-                          n.chains=2,
-                          n.adapt=1000)
-```
-
-# Q1.
-
-Here I am confused since dunif is what I found correct [here](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf).
-
-# Q2. 
-When I tried the other package and function, I got an error that object "NA" not found as below.
-
-This is using jags.model function from runjags package.
-```{r another try:run.jags}
-jags.simple <- run.jags(file='basic_model.txt',
-                          data=dat,
-                          monitor =c('u_t'),
-                          n.chains=2,
-                          n.adapt=1000)
-```
-```
+## Reference 
+[first tutorial](https://www.youtube.com/watch?v=F9jxvhXc_1w)  
+[second tutorial](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html)   
+[JAGS manual](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf)   
+[error handling guide](https://www4.stat.ncsu.edu/~reich/ABA/code/errors)

### 2020-02-16 Rcodes/JAGS_overdoses.Rmd
pdf file added

@@
 date: '2020-02-08'
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---

### 2020-02-16 Rcodes/JAGS_overdoses.Rmd
project folders rearranged

--- 2020-02-16 Rcodes/JAGS_overdoses.Rmd
@@
 $$p(p_A) \sim Beta(1,1)$$
+
 ### Overdose Model  
-
 Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**

### 2020-02-16 latex/project_goal.tex
project folders rearranged

\documentclass[]{article}

%opening
\title{Research Project}
\author{Hyeongcheol Park}

\begin{document}

\maketitle

\section{The Goal of the Project}

\begin{enumerate}

	\item{Project Goal: scrapped from e mails}

	This project is about the area of Bayesian evidence synthesis methodology oriented toward applications in public health and epidemiology.  The project will involve one or both of the following lines of enquiry:

	1. Developing new evidence synthesis models to address specific public health challenges.

	2. Developing theoretical understanding of what a priori assumptions are required to yield partial or full identification of target quantities in prototype evidence synthesis models.

	Mike builds Bayesian evidence synthesis models for public health and epidemiology applications. Typically in public health many aspects of the system of study remain hidden and surveillance data is derived from downstream effects. (e.g. think of a disease that is symptomless where only diagnoses are observed) These models also require flexibility including incorporating dynamic rates and geographic variation through the use of hierarchical components. Some examples include:

	1. Assessing the impact of the take home naloxone program on the number of overdose deaths in BC.

	2. Estimating the number of overdoses occurring within a given month and region based on ambulance call-outs, hospitalizations and other surveillance data.

	3. Estimating the size of the population with substance use disorder, opioid use disorder, individuals who inject opioids etc.

	4. The prevalence of sexually-transmitted infections based on diagnosis and testing data, combined with sexual behaviour survey data.

	Generally there are scientifically relevant questions of "information flow" in such models, very roughly put: how good do the inputs needs to be in order to get useful outputs. And the inputs could be multiple datasets plus multiple prior distributions.

	An example of this can be seen when estimating the total number of overdoses occurring based on different data sources. If an individual overdoses there is some associated probability that it is witnessed by either a passerby or emergency services. There’s a further probability an ambulance may be called or the individual is transported to hospital. Data on ambulance-attended overdoses, ED visits, and other uses of intervention each provide different information about this underlying process and can help to refine the total estimate.

	More specifically, it may not be obvious how the width of the posterior distribution on a given target parameter depends on widths of the prior distributions on various parameters plus the amount of data in each data source. But it would be scientifically relevant to know this. For example, estimating populations at risk are required for service planning. An estimate with large uncertainty (wide posterior width) would most likely not be useful.

	Of course one way to address this question is in a simulation context. Just try changing the sample sizes and the prior specification, and see what happens. And this project would likely involve some of this.

	But hopefully though there is also scope to get some clean and general answers at the cost of working with stripped-down, simplified versions of models.

	A specific avenue to investigate is whether any simplified models are amenable to the sort of "partial identification" analysis that Paul has used in other contexts. Likely the most accessible intro to this is Chapters 1 and 2 of Paul's recent book:

	Another aspect of this project can be to develop model assessment, validation, and selection within the context of partial identification. Previous work within the field of public health has used information criteria such as the Deviance Information Criterion to assess model fit and select between candidate models. Understanding the evolution of these criteria under different data sources, increased number of observations, and changes to the structure of the priors would also be valuable to investigate.

	The project is at a very early stage, quite nascent and amorphous. So the RA work is on trying to flesh out more specific research plans. For instance, can we identify which of Mike's models, and with what simplifications, are good candidates for further study?

	\item{Scratch Paragraphs: initial writing attempts}
	We have certain goals

	1. developing new evidence synthesis models for a certain public health challenges

	2. developing theorietical understanding of what a priori assumptions are required to yield identification of target quantities

	3. Estimate some possible variables of interest:
	The number of overdoses, the number of infected people and so on.

	4. Information flow: how good should the input data be in order to have fairly good enough output?

	5. how priors and data set affect the posteriors of interest?

	\begin{itemize}
		\item 5.a. by simulation in complicated context
		\item  5.b. in simplified model, some general understanding mathmatically.
	\end{itemize}

	6. Model assessment, selection and validation:
	Paul's partially identified model application

	7. For further study, which of Mike's models are good with some simplifications?

\end{enumerate}

\end{document}

### 2020-02-16 questions.R

@@ -1,14 +1,0 @@
-exp(0.028)
-
-exp(0.028)/(1+exp(0.028))
-
-p=exp(0.028)/(1+exp(0.028))
-
-
-p0 = 3/39
-p1= 3/38
-
-log(p1/(1-p1)); log((1-p1)/p1)
-log(p0/(1-p0))
-
-log((p1/(1-p1))/(p0/(1-p0)))

### 2020-02-16 report_revised_2.py

@@ -1,481 +1,0 @@
-#!/usr/bin/env python
-# coding: utf-8
-
-# # Basic Model: Ambulance + Overdose
-# 
-# 
-# 
-# ## Ambulance Call-outs Model
-# 
-# $n_{A}$: sample size   
-# $x_{A}$: the total number who confirmed they did call an ambulance  
-# $p_{A}$: probability of a person call an ambulance   
-# 
-# $$x_{A} \sim Bin(n_{A},p_{A})$$
-# 
-# **We assume $n_{A}=1000, p_A = 0.8$.**  
-# **Suppose the prior of $p_A$ is noninformative.** 
-# $$p(p_A) \sim Beta(1,1)$$
-# 
-# ## Overdose Model
-# Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**
-# Also, we have priors.
-# 
-# 
-# $$z_{t} \sim N(\mu, \sigma^{2})$$  
-# $$\lambda_{t}^{OD} = \exp(z_{t})$$  
-# $$O_{t} \sim Poi(\lambda_{t}^{OD}N)$$  
-# $$U_t \sim Bin(O_t, p_A)$$  
-# For simplicity we set N =10000 for now. We need to generate reasonable $U_t$ values first. Note that $U_t$ comes from $\mu, \sigma$ following all the way through the overdose model.  
-# 
-# 
-# 
-# $\mu=\log0.05, \sigma=1, N=10000$.   
-# We suppose survey data exists: ($n_A, x_A$) known.  
-# 
-# **We set for our prior parameters:** $$\mu \sim U(-10,0)$$  $$\sigma \sim U(0,5)$$
-
-# In[1]:
-
-
-import pandas as pd
-import numpy as np
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-import pymc3 as pm
-import theano.tensor as tt
-import seaborn as sns
-
-
-# In[2]:
-
-
-import platform
-platform.python_version()
-
-
-# Below is the function that generate data from simulation.
-
-# In[3]:
-
-
-def sim_ambulance_overdoses(random=1, mu_z=np.log(0.05), sigma_z=1.,
-                            N=10000, p_a=0.8, n_a=1000, n_T=12):
-    '''
-    Simulate overdoses data, ambulance-call outs survey data and related variables 
-    -----------
-    mu_z- float 
-        log rate for lambda
-    
-    sigma_z -float
-        variance of log lambda
-    
-    N - int
-        Total size of population
[…]

### 2020-02-19 Rcodes/JAGS_overdoses.Rmd
RJAGS output done

@@
 # install packages
-if (!require(runjags)) install.packages("runjags")
-if (!require(rjags)) install.packages("rjags")
-if (!require(coda)) install.packages("coda")
+if (!require(rjags)) install.packages("rjags", dependencies = TRUE)
+if (!require(coda)) install.packages("coda", dependencies = TRUE)
+if (!require(tidyverse)) install.packages("tidyverse", dependencies = TRUE)
+if (!require(tinytex)) install.packages("tinytex", dependencies = TRUE)
 
-library('runjags')
 library('rjags')
 library('coda')
+library('tidyverse')
+library('tinytex')
+
 ```
@@
 ``` {r run the first model without error }
-jags.simple <- jags.model(file='basic_model.txt',
-                          data=dat)
+interations = 1000
+burnin= floor(interations/2)
+chains=2
+# inits = list()
+simple.model <- jags.model(file='basic_model.txt',
+                          data=dat,
+                          n.chains = chains)
+simple.model
+```
+
+### O_t
+
+```{r}
+params= c('o_t')
+samples <- coda.samples(simple.model, params, n.iter = 1000)
+```
+
+# Q1: is init necessary? and it's a characteristic of Gibbs sampling?
+
+```{r }
+summary(window(samples), start=burnin)
+```
+
+```{r include=FALSE}
+plot(samples)
+```
+
+### Boxplots of O_t
+
+# q2: I see two elements from the samples list. Which one I should use it or should I use both?
+
+```{r yield a data frame}
+temp = as.matrix(samples)
+colnames(temp) <- seq(1,12)
+head(temp)
+df_o_t <- as.data.frame(temp)
+head(df_o_t)
+```
+
+```{r tidy the data frame}
+df_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
+df_o_t$month <- factor(df_o_t$month,levels = seq(1,12))
+str(df_o_t)
+```
+
+```{r}
+o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
+o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
+
+ggplot()+geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')+geom_boxplot(aes(x=month,y=o_t),data = df_o_t)
+```
+
+```{r}
+params= c('u_t','x_a')
+ppc <- coda.samples(simple.model, params, n.iter = 1000)
+```
+
+```{r}
+summary(window(ppc),start=burnin)
+
+```
+
+```{r eval=FALSE, include=FALSE}
+plot(ppc)
+```
+
+```{r}
+temp = as.matrix(ppc)
+
+df_u_t=temp[,1:12]
+df_x_a=temp[,13:24]
+
+test <- as.matrix(df_u_t)
+colnames(test) <- seq(1,12)
+head(test)
+mtx_u_t <- as.data.frame(test)
+head(mtx_u_t)
+
+df_u_t <- gather(mtx_u_t, key = 'month',value = 'u_t')
+head(df_u_t)
+
+df_u_t$month <- factor(df_u_t$month,levels = seq(1,12))
+str(df_u_t)
+
+u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
+u_t_values$month <- factor(u_t_values$month,levels = seq(1,12))
+
+ggplot()+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)
+```
+
+```{r}
+
+temp = as.matrix(ppc)
+df_x_a=temp[,13:24]
+
+test <- as.matrix(df_x_a)
+colnames(test) <- seq(1,12)
+head(test)
+mtx_x_a <- as.data.frame(test)
+head(mtx_x_a)
+
+df_x_a <- gather(mtx_x_a, key = 'month',value = 'x_a')
+head(df_x_a)
+
+df_x_a$month <- factor(df_x_a$month,levels = seq(1,12))
+str(df_x_a)
+
+x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
+x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
+
+ggplot()+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')+geom_boxplot(aes(x=month,y=x_a),data = df_x_a)
 ```

### 2020-02-19 Rcodes/JAGS_overdoses.Rmd
trimmed the rmd

@@
 ```
+### Predictive Posterior Checks
 
@@
 ```
+### $U_t$: Predictive Posterior Checks
 
@@
 colnames(test) <- seq(1,12)
-head(test)
 mtx_u_t <- as.data.frame(test)
-head(mtx_u_t)
 
 df_u_t <- gather(mtx_u_t, key = 'month',value = 'u_t')
-head(df_u_t)
 
 df_u_t$month <- factor(df_u_t$month,levels = seq(1,12))
-str(df_u_t)
 
@@
 ```
+
+### $x_A$: Predictive Posterior Checks
 
@@
 colnames(test) <- seq(1,12)
-head(test)
+
 mtx_x_a <- as.data.frame(test)
-head(mtx_x_a)
 
 df_x_a <- gather(mtx_x_a, key = 'month',value = 'x_a')
-head(df_x_a)
 
 df_x_a$month <- factor(df_x_a$month,levels = seq(1,12))
-str(df_x_a)
 

### 2020-03-02 Rcodes/JAGS_overdoses.Rmd
flu over and found where I did it wrong !!!

@@
 p_a ~ dbeta(alpha, beta)
-mu ~ dunif(mu_a, mu_b)
-sigma ~ dunif(sigma_a, sigma_b)
+mu_z ~ dunif(mu_a, mu_b)
+sigma_z ~ dunif(sigma_a, sigma_b)
+
+# unify variable names: 
+# z_t <- z lmb_t <- lambda
+# length unification: z_t is length 12 vector
+
+ ## ambulance model
+for (i in 1:n) {
+  #Likelihood
+  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
+}
+
+# overdose 
+for (i in 1:n) {
 
   ## the latent variables
-z ~ dnorm(mu, 1/(sigma^2))
-lambda <- exp(z)
+  z_t[i]~ dnorm(mu_z, 1/(sigma_z^2))
+  lmb_t[i] <- exp(z_t[i])
 
-for (i in 1:n) {
-  ## ambulance model 
-  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
-}
-for (i in 1:n) {
   ## overdose model
-  o_t[i] ~ dpois(lambda*N) # total overdoses per month
-}
-for (i in 1:n) {
+  o_t[i] ~ dpois(lmb_t[i]*N) # total overdoses per month
+# Note that from pymc3 gamma was used instead of Pois dist
   u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
@@
             # likelihood 
-            'u_t'=u_t, 
-            'x_a'=x_a,  
-            'N'=N, # the population
-            'n'= n, # total months
-            'n_a'=n_a
+            'u_t'=u_t, # giving data 
+            'x_a'=x_a,   # giving data
+            'N'=N, # the population 10000
+            'n'= n, # total months 12
+            'n_a'=n_a # survey size 1000
 
@@
 ``` {r run the first model without error }
-interations = 1000
-burnin= floor(interations/2)
 chains=2
@@
 samples <- coda.samples(simple.model, params, n.iter = 1000)
+# guess it's getting posterior samples ?
 ```
 
-# Q1: is init necessary? and it's a characteristic of Gibbs sampling?
-
 ```{r }
+interations = 1000
+burnin= floor(interations/2)
 summary(window(samples), start=burnin)
@@
 ```
+# q1: what is the equivalent plot that we can see we have enough iteration?
 
@@
 colnames(temp) <- seq(1,12)
-head(temp)
 df_o_t <- as.data.frame(temp)
@@
 ```{r tidy the data frame}
-df_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
-df_o_t$month <- factor(df_o_t$month,levels = seq(1,12))
-str(df_o_t)
+trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
+trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
+str(trace_o_t)
 ```
@@
 ```{r}
+# real values of the data set from pymc3 samples
 o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
+# factorizing the months for box plot visualization
 o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
 
-ggplot()+geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')+geom_boxplot(aes(x=month,y=o_t),data = df_o_t)
+ggplot()+
+  # boxplot from the trace 
+  geom_boxplot(aes(x=month,y=o_t), color='grey',alpha = 0.01,data = trace_o_t)+
+  # real values as red dots
+  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
+
 ```
+
 ### Predictive Posterior Checks
@@
 
-ggplot()+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)
+ggplot()+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
 ```

### 2020-03-04 Rcodes/JAGS_overdoses.Rmd
almost done with JAGS and flu shot is important

@@
 ---
-## Basic Model: Ambulance + Overdose
-### Ambulance Call-outs Model
+# Basic Model: Ambulance + Overdose
+## Ambulance Call-outs Model
 $n_{A}$: sample size   
@@
 
-### Overdose Model  
+## Overdose Model  
 Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**
@@
 ## define the priors
+
 p_a ~ dbeta(alpha, beta)
@@
 sigma_z ~ dunif(sigma_a, sigma_b)
-
-# unify variable names: 
-# z_t <- z lmb_t <- lambda
-# length unification: z_t is length 12 vector
 
@@
   x_a[i] ~ dbin(p_a, n_a) # each survey result for month
+
 }
@@
 ```{r Defining Values}
-n <-  length(df$o_t) # we have 12 samples
+n_T <-  length(df$o_t) # we have 12 samples
 n_a <- 1000
@@
             'N'=N, # the population 10000
-            'n'= n, # total months 12
+            'n'= n_T, # total months 12
             'n_a'=n_a # survey size 1000
@@
 **Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
-``` {r run the first model without error }
+``` {r run the first model}
 chains=2
@@
 ```{r}
-params= c('o_t')
+params= c('o_t','p_a')
 samples <- coda.samples(simple.model, params, n.iter = 1000)
@@
 ```{r yield a data frame}
-temp = as.matrix(samples)
+
+pst_mtx = as.matrix(samples)
+temp = pst_mtx[,1:12]
+p_a <- pst_mtx[,13]
+head(temp)
+length(p_a)
 colnames(temp) <- seq(1,12)
@@
 head(df_o_t)
+str(df_o_t)
 ```
@@
 str(trace_o_t)
+head(trace_o_t,n = 24)
 ```
@@
   # boxplot from the trace 
-  geom_boxplot(aes(x=month,y=o_t), color='grey',alpha = 0.01,data = trace_o_t)+
+  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
   # real values as red dots
@@
 
-### Predictive Posterior Checks
+## Predictive Posterior Checks
 
 ```{r}
-params= c('u_t','x_a')
-ppc <- coda.samples(simple.model, params, n.iter = 1000)
+df_u_t <- data.frame()
+
+for (i in 1:m) {
+  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
+  df_u_t=rbind(df_u_t,obs)
+}
+colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
+str(df_u_t)
+```
+
+### $U_t$: Predictive Posterior Checks
+
+```{r}
+ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
+summary(ppc_u_t)
+u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
+# u_t_values$month <- factor(u_t_values$month,levels = seq(1,12))
+ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
+
+ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
+```
+
+# todo: finish this part.
+### $x_A$: Predictive Posterior Checks
+
+```{r Generate posterior predictive samples of $x_a$}
+
+df_x_a <- vector()
+for (i in 1:n_T) {
+  obs <- rbinom(m,n_a,p_a)
+  df_x_a <- cbind(df_x_a,obs)
+}
+
+head(df_x_a)
+colnames(df_x_a) <- seq(1:12)
 ```
@@
 ```{r}
-summary(window(ppc),start=burnin)
 
-```
+df_x_a <- as.data.frame(df_x_a)
+ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
 
-```{r eval=FALSE, include=FALSE}
-plot(ppc)
-```
-### $U_t$: Predictive Posterior Checks
-
-```{r}
-temp = as.matrix(ppc)
-
-df_u_t=temp[,1:12]
-df_x_a=temp[,13:24]
-
-test <- as.matrix(df_u_t)
-colnames(test) <- seq(1,12)
-mtx_u_t <- as.data.frame(test)
-
-df_u_t <- gather(mtx_u_t, key = 'month',value = 'u_t')
-
-df_u_t$month <- factor(df_u_t$month,levels = seq(1,12))
-
-u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
-u_t_values$month <- factor(u_t_values$month,levels = seq(1,12))
-
-ggplot()+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
-```
-
-### $x_A$: Predictive Posterior Checks
-
-```{r}
-
-temp = as.matrix(ppc)
-df_x_a=temp[,13:24]
-
-test <- as.matrix(df_x_a)
-colnames(test) <- seq(1,12)
-
-mtx_x_a <- as.data.frame(test)
-
-df_x_a <- gather(mtx_x_a, key = 'month',value = 'x_a')
-
-df_x_a$month <- factor(df_x_a$month,levels = seq(1,12))
+ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
 
@@
 
-ggplot()+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')+geom_boxplot(aes(x=month,y=x_a),data = df_x_a)
+ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')
+```
+
+# Contamination of $p_A$
+
+Now, suppose the survey data gives us a wrong (biased) $p_A$ value.
+
+Bias = $\theta -\hat{\theta} = p_A- \hat{p}_A$   
+$\hat{p}_A = p_A + bias(p_A)$
+
+Three more data sets are given: unbiased, overestimated, underestimated $p_A$.
+
+```{r}
+## write a function that led to compare o_t, u_t and x_a. 
+#first we need a fuction that gives us data, model, trace, and ppc.
+test_robust <- function(file=file, random=1, N=10000, p_a=0.8, bias = -0.2, n_a=1000, n_T=12) {
+  df <- read.csv(file = file)
+  df$X <- NULL 
+  df$month <- seq(1:12)
+# obtain the (biased) data 
+  dat <- list(
+            # priors for ambulance model
+            'alpha' = 1,
+            'beta' =  1,
+
+            # priors for overdose model 
+            'mu_a'=(-10),
+            'mu_b'=0,
+            'sigma_a'=0,
+            'sigma_b'=5,
+
+            # likelihood 
+            'u_t'=df$u_t, # giving data 
+            'x_a'=df$x_a,   # giving data
+            'N'=N, # the population 10000
+            'n'= n, # total months 12
+            'n_a'=n_a # survey size 1000
+
+            )
+# run the model 
+  chains=2
+  # target 1 to save
+  simple.model <- jags.model(file='basic_model.txt',
+                          data=dat,
+                          n.chains = chains)
+
+# get the samples of O_t, p_a
+  params= c('o_t','p_a')
+  # target 2 to save
+  samples <- coda.samples(simple.model, params, n.iter = 1000)
+
+  # tidy p_a: target 3 
+  p_a <- pst_mtx[,13]
+
+  ## tidy o_t
+  pst_mtx = as.matrix(samples)
+  temp = pst_mtx[,1:12]
+  colnames(temp) <- seq(1,12)
+  df_o_t <- as.data.frame(temp)
+  trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
+  trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
+  # real values of the data set from pymc3 samples
+  # o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
+  # # factorizing the months for box plot visualization
+  # o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
+  # boxplots of o_t
+
+  # tidy u_t pp samples
+  df_u_t <- data.frame()
+  for (i in 1:m) {
+   obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
+   df_u_t=rbind(df_u_t,obs)
+  }
+  colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
+  ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
+  ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
+  # u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
+
+  ## tidy x_a
+  df_x_a <- vector() 
+  for (i in 1:n) {
+    obs <- rbinom(m,n_a,p_a)
+    df_x_a <- cbind(df_x_a,obs)
+  }
+  colnames(df_x_a) <- seq(1:12)
+  df_x_a <- as.data.frame(df_x_a)
+  ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
+  ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
+  # x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
+  # x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
+  ppc = list('u_t'=ppc_u_t, 'x_a'=ppc_x_a)
+  mylist=list('data'=df,'model'=simple.model,'trace'=trace_o_t,'ppc'=ppc)
+  return (mylist)
+}
+
+# then visualization function which gives us three different types of boxplots.
+```
+
+```{r}
+my_list_unbiased = test_robust(file = './basic_data.csv', bias =0)
+my_list_under = test_robust(file = './under_p_a_data.csv',bias=-0.2)
+my_list_over = test_robust(file='./over_p_a_data.csv',bias = +0.1)
+```
+
+```{r Check whether we use the same data set}
+print(my_list_unbiased$data$u_t)
+print(my_list_under$data$u_t)
+print(my_list_over$data$u_t)
+
+print(my_list_unbiased$data$x_a)
+print(my_list_under$data$x_a)
+print(my_list_over$data$x_a)
+```
+```{r}
+ head(my_list_unbiased$ppc$u_t)
+```
+
+```{r}
+# finish this function
+visualization <- function(mylist= None, post= F, u_t = F, x_a = F, string='string') {
+  data= mylist$data
+  if (post == T) {
+    # boxplots of o_t
+    trace_o_t = mylist$trace
+  p <- ggplot()+
+  # boxplot from the trace 
+  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
+  # real values as red dots
+  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
+
+  }
+
+  if (u_t == T) {
+    ppc_u_t = mylist$ppc$u_t
+    # boxplots of u_t
+  p <- ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=data$month, y=data$u_t),color='red')
+
+  }
+
+  if (x_a == T) {
+    p <- ppc_x_a = mylist$ppc$x_a
+    ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=data$month, y=data$x_a),color='red')
+  }
+  return(p)
+}
+```
+
+```{r}
+visualization(my_list_unbiased, post= T)
+```
+
+# Trim here and make a function and give us the plots.
+```{r}
+## Trim here and make a function and give us the plots.
+df <- read.csv('./basic_data.csv')
+df$X <- NULL 
+# obtain the (biased) data 
+dat <- list(
+            # priors for ambulance model
+            'alpha' = 1,
+            'beta' =  1,
+
+            # priors for overdose model 
+            'mu_a'=(-10),
+            'mu_b'=0,
+            'sigma_a'=0,
+            'sigma_b'=5,
+
+            # likelihood 
+            'u_t'=df$u_t, # giving data 
+            'x_a'=df$x_a,   # giving data
+            'N'=N, # the population 10000
+            'n'= n, # total months 12
+            'n_a'=n_a # survey size 1000
+
+            )
+# run the model 
+chains=2
+simple.model <- jags.model(file='basic_model.txt',
+                          data=dat,
+                          n.chains = chains)
+
+# get the samples of O_t, p_a
+params= c('o_t','p_a')
+samples <- coda.samples(simple.model, params, n.iter = 1000)
+
+# tidy p_a
+p_a <- pst_mtx[,13]
+
+## tidy o_t
+pst_mtx = as.matrix(samples)
+temp = pst_mtx[,1:12]
+colnames(temp) <- seq(1,12)
+df_o_t <- as.data.frame(temp)
+trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
+trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
+# real values of the data set from pymc3 samples
+o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
+# factorizing the months for box plot visualization
+o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
+# boxplots of o_t
+ggplot()+
+  # boxplot from the trace 
+  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
+  # real values as red dots
+  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
+
+# tidy u_t pp samples
+df_u_t <- data.frame()
+for (i in 1:m) {
+  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
+  df_u_t=rbind(df_u_t,obs)
+}
+colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
+ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
+ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
+u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
+# boxplots of u_t
+ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
+
+## tidy x_a
+df_x_a <- vector() 
+for (i in 1:n) {
+  obs <- rbinom(m,n_a,p_a)
+  df_x_a <- cbind(df_x_a,obs)
+}
+colnames(df_x_a) <- seq(1:12)
+df_x_a <- as.data.frame(df_x_a)
+ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
+ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
+x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
+x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
+
+ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')
 ```

### 2020-03-04 latex/newpaper.tex
almost done with JAGS and flu shot is important

--- 2020-02-07 article_4/newpaper.tex
@@
 The opioid crisis is a major issue in North America including British Columbia, Canada.
-British Columbia is in the midst of a drug overdose crisis due to illicit opioids and the circumstance is getting worse; According to the BC Coroners Service, there were more than 930 apparent illicit drug overdose deaths in BC from Jan.
+British Columbia is in the midst of a drug overdose crisis due to illicit opioids and the circumstances are getting worse.
+According to the BC Coroners Service, there were more than 930 apparent illicit drug overdose deaths in BC from Jan.
 1 to Dec.
@@
 This compares to 518 in 2015, an increase of 79.2\%.
-\cite{bccdc_opioid}  The goal of this project is to apply Bayesian evidence synthesis to understand better the opoid crisis in British Columbia, Canada.
+\cite{bccdc_opioid}  The goal of this project is to apply Bayesian evidence synthesis to better understand the opoid crisis in British Columbia, Canada.
 \\
 
-One difficulty to cope with the opioid crisis is that the total number of overdoses is unknown.
-The reason is that the use of prescription opioids often leads to use illicit opioids and the illicit usage cannot be known.
+One difficulty in coping with the opioid crisis is that the total number of overdoses is unknown.
+The reason is that the use of prescription opioids often leads to use of illicit opioids and the illicit usage cannot be known.
 Many of those who become addicted to opioids do so after initially receiving a prescription.
@@
 In fact, about 80\% of people using heroin started with a prescription to another opioid.
-After using heroin, however, 23\% of individuals develop opioid addiction.\cite{opioid_desc} [attempt: need to give a backstory-why isn't this observable?
-The number of overdoses]\\
+After using heroin, however, 23\% of individuals develop opioid addiction.\cite{opioid_desc} \\
 
-Since the total overdoses are unknown, the number needs to be estimated.
-Bayesian statistics can be a way to approach the problem and give us the good estimate of the number.
+Since the total number of overdoses is unknown, this number needs to be estimated.
+Bayesian statistics can be a way to approach the problem and give us a good estimate of the number [\textbf{defensible?
+depending on the data, a precise estimate may not be possible}].
 Bayesian statistics is a theory in the field of statistics based on the Bayesian interpretation of probability where probability expresses a degree of belief in an event.
@@
 
-\section{methods}
+\section{Methods}
 
@@
 
-The number of overdoses is our ultimate interest of estimation.
+The number of overdoses is our ultimate target of estimation.
 To achive the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
-The posterior belief is coming from both parts of the information.
-Following equation express the idea in a mathmatical form:
+The posterior belief is coming from both of these sources of the information.
+The following equation express the idea in a mathmatical form:
 \begin{equation}
@@
 \left.\begin{aligned}
-p(\theta| y ) = \frac{ p(\theta, y)   }{ p(y)} = \frac{ p(\theta) p(y| \theta)}{ p(y)} \propto p(\theta) p(y| \theta) \end{aligned}\right.
+p(O_t| y ) = \frac{ p(O_t, y)   }{ p(y)} = \frac{ p(O_t) p(y| O_t)}{ p(y)} \propto p(O_t) p(y| O_t) \end{aligned}\right.
 \end{equation}
-where $\theta$ represents the target of estimation, the total overdoses and $y$ represents the collected data set providing relevant samples of the target variable.
-In Bayesian concepts, the variable of interest $\theta$ is a fixed number but the number is yielded from a certain distribution; That is, $\theta$ is considered to be a sample of a random variable.
+where $O_t$ is the total number of overdoses in a given month $t$ in British Columbia, and $y$ represents the collected data set providing relevant samples of the target variable.
+In Bayesian concepts, the variable of interest $O_t$ is a fixed number but the number is yielded from a certain distribution where the distribution describes knowledge about $O_t$; That is, $O_t$ is considered to be a sample of a random variable.
 The problem, however, is that the data $y$ is not available.
-We can have our prior belief about the total overdoses but it is not possible to obtain any data.
 Hence it is needed to approach the estimation in a indirect fashion.
 Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
-[attempt: Give a graphical model representation?
-state that(?)]
+\\
 \begin{figure}[htb]
@@
 	\includegraphics[width=0.5\linewidth]{Figures/simple_drawing}
-	\caption{Example of estimating the total overdoses indirectly given two data sets: ambulance attended overdoses and survey data. N is the population size}
+	\caption{Example of estimating the total overdoses indirectly given two data sets: ambulance attended overdoses and survey data.}
 	\label{simple_draw}
 \end{figure}
-\\
 
- Let $O_t$ the number of overdoses in a given month $t$ in British Columbia.
-[attempt: specify a geographic area] Suppose there was a survey  conducted to estimate the proportion of ambulance call $p_A$ among the overdoses.
-$p_A$ is assumed constant across time for simplicity.
-Let $n_{A}$ the sample size of the survey and $x_{A}$ to be the total number who confirmed they did call ambulance.
-It is assumed that $x_{A}$ follows a Binomial distribution:
-
-\begin{equation}
-\label{ambulance}
-\left.\begin{aligned}
-x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
-\text{ambulance call-outs model}
-\end{equation}
-
-The total overdoses need to be modeled.
-The simplest conceptual model is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
-\cite{Irvine:modelling} Denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$.
-It is assumed that the total overdose $O_t$ follows Poisson distribution where the population of the region of interest is $N$.
+ The simplest conceptual model described in Figure \ref{simple_draw} is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
+\cite{Irvine:modelling} Then denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$.
+It is assumed that the total overdose $O_t$ follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
+Note that the population size, N, is unknown and vary over time in reality but it is assumed known constant for the investigator here for simplicity.
 
@@
 
-Estimation of $O_t$ is not straightforward since none of the variables ($\mu$, $\sigma$, $N$) determining $O_t$  is known.
-Hence $O_t$ should be inferred from using $U_t$ and \(p_A\), where $p_A$ is the ambulance call out rate and \(U_t\) is the number of  ambulance-attended overdoses at a time point $t$.
+Since estimation of $O_t$ is not straightforward as none of the variables ($\mu$, $\sigma$, $N$) determining $O_t$  is known, we assume there was a survey conducted to infer $O_t$.
+The survey asked witnesses of opioid overdoses in danger whether someone called an ambulance.
+This provides estimate of the proportion of ambulance call $p_A$ among the overdoses.
+Let $n_{A}$ the sample size of the survey and $x_{A}$ to be the total number who confirmed they did call ambulance.
+It is assumed that $x_{A}$ follows a Binomial distribution:
+
+\begin{equation}
+\label{ambulance}
+\left.\begin{aligned}
+x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
+\text{ambulance call-outs model}
+\end{equation}
+
+Let \(U_t\) the number of  ambulance-attended overdoses at a time point $t$.
 In general, the data of ambulance-attended overdoses \(U_t\) can be obtained.
@@
 
-Now $O_t$ can be estimated as $p_A$ can be infered by survey data and the data regarding $U_t$ is given.
-We suggest a simple model as a start where the model only combines Ambulance Call-outs Model (\ref{ambulance}) and Overdose Model (\ref{overdose}).
-
-The next step is to run some simulations to figure out how different types of inputs lead some changes of output. To do so, the simple model is illustrated below.\\
+Note that now $O_t$ can be estimated because $p_A$ can be infered from the survey data and the data regarding $U_t$ is given.
+For the moment $p_A$ is assumed constant across time for simplicity.
+We suggest a simple model as a start where the model only combines Overdose Model (\ref{overdose}) and Ambulance Call-outs Model (\ref{ambulance}) with the relation (\ref{over_amb}).
+The next step is to run some simulations to figure out how different types of inputs lead some changes of output.
+\\
 
@@
 The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.
-Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity as in figure \ref{simple_draw}.
 The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality.

### 2020-03-04 Rcodes/JAGS_overdoses.Rmd
JAGS done: found an finding: x_a ppc and u_t ppc different

@@
   # target 2 to save
-  samples <- coda.samples(simple.model, params, n.iter = 1000)
+  samples <- coda.samples(simple.model, params, n.iter = 2000)
 
@@
   trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
-  # real values of the data set from pymc3 samples
-  # o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
-  # # factorizing the months for box plot visualization
-  # o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
-  # boxplots of o_t
 
@@
   ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
-  # u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
 
@@
   ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
-  # x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
-  # x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
   ppc = list('u_t'=ppc_u_t, 'x_a'=ppc_x_a)
@@
   if (x_a == T) {
-    p <- ppc_x_a = mylist$ppc$x_a
-    ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=data$month, y=data$x_a),color='red')
+    ppc_x_a = mylist$ppc$x_a
+    p <- ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=data$month, y=data$x_a),color='red')
   }
@@
 ```{r}
-visualization(my_list_unbiased, post= T)
+visualization(my_list_unbiased, post=T)
+visualization(my_list_under,post= T)
+visualization(my_list_over,post=T)
+```
+
+```{r}
+visualization(my_list_unbiased, u_t=T)
+visualization(my_list_under,u_t= T)
+visualization(my_list_over,u_t=T)
+```
+
+```{r}
+visualization(my_list_unbiased, x_a=T)
+visualization(my_list_under,x_a= T)
+visualization(my_list_over,x_a=T)
 ```
@@
 # Trim here and make a function and give us the plots.
-```{r}
+```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 ## Trim here and make a function and give us the plots.

### 2020-03-04 Rcodes/JAGS_overdoses.Rmd
pdf generated

@@
 ```{r Defining Values}
-n_T <-  length(df$o_t) # we have 12 samples
+n_T <- length(df$o_t)
 n_a <- 1000
@@
 ```{r include=FALSE}
-plot(samples)
+# plot(samples)
 ```
@@
 head(df_o_t)
-str(df_o_t)
+
 ```
@@
 df_u_t <- data.frame()
-
+m=length(p_a)
 for (i in 1:m) {
@@
   df_x_a <- vector() 
-  for (i in 1:n) {
+  for (i in 1:n_T) {
     obs <- rbinom(m,n_a,p_a)

### 2020-03-04 Rcodes/JAGS_overdoses.Rmd
JAGS DONE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@
   samples <- coda.samples(simple.model, params, n.iter = 2000)
-
+  pst_mtx = as.matrix(samples)
   # tidy p_a: target 3 
@@
   ## tidy o_t
-  pst_mtx = as.matrix(samples)
   temp = pst_mtx[,1:12]
@@
   df_u_t <- data.frame()
+  m = length(p_a)
   for (i in 1:m) {
@@
 ```
+############# Below here is for the purpose of debugging ##################
 
-# Trim here and make a function and give us the plots.
 ```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 ## Trim here and make a function and give us the plots.
-df <- read.csv('./basic_data.csv')
+df <- read.csv('./under_p_a_data.csv')
 df$X <- NULL 
+df$bias <- NULL
 # obtain the (biased) data 
@@
 # run the model 
+head(df)
+```
+
+```{r}
 chains=2
-simple.model <- jags.model(file='basic_model.txt',
+under.model <- jags.model(file='basic_model.txt',
                           data=dat,
@@
 
+```
+
+```{r}
+# inits = list()
+
 # get the samples of O_t, p_a
 params= c('o_t','p_a')
-samples <- coda.samples(simple.model, params, n.iter = 1000)
-
+samples <- coda.samples(under.model, params, n.iter = 1000)
+pst_mtx = as.matrix(samples)
+head(pst_mtx)
 # tidy p_a
 p_a <- pst_mtx[,13]
+head(p_a)
+```
+
+```{r}
 
@@
   geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
+```
+
+```{r}
 
@@
 ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
-
+```
+```{r}
 ## tidy x_a
+m=length(p_a)
 df_x_a <- vector() 
-for (i in 1:n) {
+for (i in 1:n_T) {
   obs <- rbinom(m,n_a,p_a)
@@
 }
+
 colnames(df_x_a) <- seq(1:12)

### 2020-03-11 Rcodes/JAGS_overdoses.Rmd
N uncertain work -ing

@@
 ```
-############# Below here is for the purpose of debugging ##################
+# Uncertainty of N
+Now we suppose the population size of BC, N, is unknown.
+We may have some information about the population size and transform it into a prior distribution.
+Let's assume from the point that we don't have a good understanding of the sample size.
+But N should be at least bigger than the sample size of survey, the ambulance attended overdoses.
+We have 12 data points of $U_t$.
+We also have 12 data points of $x_A$.
 
-```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-## Trim here and make a function and give us the plots.
-df <- read.csv('./under_p_a_data.csv')
-df$X <- NULL 
-df$bias <- NULL
-# obtain the (biased) data 
-dat <- list(
+$$N \geq max(\vec{U_t}, \vec{x_A})$$
+
+Where $\vec{U_t}, \vec{x_A}$ are the vectors of $U_t$ and $x_A$.
+
+Then, we need a theoretical maximum value of N.
+BC is a province of Canada.
+For now, since I do not have a good idea about the maximum possible number of N, let's just say two times of the real value is the maximum.
+
+$$N \in  \{a, a+1, a+2, ..., b\} \\ \textbf{where } a= \max(\vec{U_t}, n_A), b= 2 \times \textbf{real value of  N}$$
+
+However, in the model file below, we will use a prior such that 
+$$N \in  \{0,1,2, ..., b\} \\ \textbf{where }  b= 2 \times \textbf{real value of  N}$$ since we don't know how to construct such prior as described below.
+
+```{r Enter the models}
+cat("model{
+## define the priors
+
+p_a ~ dbeta(alpha, beta)
+mu_z ~ dunif(mu_a, mu_b)
+sigma_z ~ dunif(sigma_a, sigma_b)
+
+# Uncertain N takes in to the model
+N ~ dcat(p_i)
+
+ ## ambulance model
+for (i in 1:n) {
+  #Likelihood
+  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
+
+}
+
+# overdose 
+for (i in 1:n) {
+
+  ## the latent variables
+  z_t[i]~ dnorm(mu_z, 1/(sigma_z^2))
+  lmb_t[i] <- exp(z_t[i])
+
+  ## overdose model
+  o_t[i] ~ dpois(lmb_t[i]*N) # total overdoses per month
+  # Note that from pymc3 gamma was used instead of Pois dist
+  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
+}
+
+}", file='N_uncertain_model.txt')
+```
+
+Pre-set variables.  
+```{r Defining Values}
+p_i <- seq(1:(2*N))
+n_T <- length(df$o_t)
+n_a <- 1000
+# N <- 10000
+u_t <- df$u_t
+x_a <- df$x_a
+```
+
+Define the list providing the values of the variables and the parameters for the priors of the model.   
+```{r from variable name to \'variable_name\'}
+dat.N <- list(# prior for N
+            'p_i' = p_i,
             # priors for ambulance model
@@
             # likelihood 
-            'u_t'=df$u_t, # giving data 
-            'x_a'=df$x_a,   # giving data
-            'N'=N, # the population 10000
+            'u_t'=u_t, # giving data 
+            'x_a'=x_a,   # giving data
             'n'= n_T, # total months 12
@@
             )
-# run the model 
-head(df)
+```
+
+``` {r run the first model}
+chains=2
+# inits = list()
+simple.N.model <- jags.model(file='N_uncertain_model.txt',
+                          data=dat.N,
+                          n.chains = chains)
+simple.N.model
 ```
@@
 ```{r}
-chains=2
-under.model <- jags.model(file='basic_model.txt',
-                          data=dat,
-                          n.chains = chains)
-
+test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 1000)
+interations = 1000
+burnin= floor(interations/2)
+summary(window(test), start=burnin)
+head(test)
 ```
 
+```{r get the samples of N and show a distribution}
+pst_mtx.N = as.matrix(test)
+head(pst_mtx.N)
+N <- pst_mtx.N[,1]
+```
 ```{r}
-# inits = list()
-
-# get the samples of O_t, p_a
-params= c('o_t','p_a')
-samples <- coda.samples(under.model, params, n.iter = 1000)
-pst_mtx = as.matrix(samples)
-head(pst_mtx)
-# tidy p_a
-p_a <- pst_mtx[,13]
-head(p_a)
+ggplot(data = as.data.frame(N), aes(x= N)) +geom_histogram()+ geom_vline(xintercept =10000)
 ```
 
+It took a lot of time to fit the bayesian model this time.
+We have overestimated N value whose mode is around 17000 while the real value is 10000.
+Unif(0,20000) has 10000 as its mean.
+I guess the b was too large.
+I will wait for the reply from Paul. 
 ```{r}
-
-## tidy o_t
-pst_mtx = as.matrix(samples)
-temp = pst_mtx[,1:12]
-colnames(temp) <- seq(1,12)
-df_o_t <- as.data.frame(temp)
-trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
-trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
-# real values of the data set from pymc3 samples
-o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
-# factorizing the months for box plot visualization
-o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
-# boxplots of o_t
-ggplot()+
-  # boxplot from the trace 
-  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
-  # real values as red dots
-  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
-```
-
-```{r}
-
-# tidy u_t pp samples
-df_u_t <- data.frame()
-for (i in 1:m) {
-  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
-  df_u_t=rbind(df_u_t,obs)
-}
-colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
-ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
-ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
-u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
-# boxplots of u_t
-ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
-```
-```{r}
-## tidy x_a
-m=length(p_a)
-df_x_a <- vector() 
-for (i in 1:n_T) {
-  obs <- rbinom(m,n_a,p_a)
-  df_x_a <- cbind(df_x_a,obs)
-}
-
-colnames(df_x_a) <- seq(1:12)
-df_x_a <- as.data.frame(df_x_a)
-ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
-ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
-x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
-x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
-
-ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')
+plot(test)
 ```

### 2020-03-11 latex/newpaper.tex
N uncertain work -ing

@@
 \normalsize 
-The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.
-The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality.
-The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.
-\\
+The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.  \\
 
@@
 There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets are simulated as follows. 
-The true value of $p_A$ was set $p_A=0.8$ for the survey data.
+The true value of $p_A$ was set as $p_A=0.8$ for the survey data. [n_A? the setup described here above has a single x ]
 It is assumed that the data was collected for a year (t=1,2,3, ..., 12) and 
@@
 The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).
-[Delete: Then the 12 sueveys were combined into a single survey since $p_A$ is assumed fixed across time in this contexts.] Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\
+Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 p(p_A) \sim Beta(1,1)
-\text{			noninformative prior of ambulance model}
+\text{			noninformative prior for ambulance model}
 \end{equation}
@@
 \end{aligned}\right\} 
-\text{			noninformative prior of overdose model}
+\text{			noninformative prior for overdose model}
 \end{equation}
@@
 
+\subsection{Variying the Assumptions}
+The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality.
+The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.
+
 \bibliographystyle{unsrt}

### 2020-03-11 latex/newpaper.tex
histogram generated

@@
 \normalsize 
-There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets are simulated as follows. 
-The true value of $p_A$ was set as $p_A=0.8$ for the survey data. [n_A? the setup described here above has a single x ]
-It is assumed that the data was collected for a year (t=1,2,3, ..., 12) and 
-$x_t$ values were independentally generated from the Binomial distribution (\ref{ambulance}).
-In terms of overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
+There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets were simulated as follows. 
+It was assumed that the two data sets were collected for a year (t=1,2,3, ..., 12).
+The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months.
+The $x_A$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
+In terms of the overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
 The vector of $O_t$ was generated following the overdose model (\ref{overdose}).
@@
 The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).
-Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\
+Note that only $U_t$ and $x_A$ are known as the likelihood and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 The result from the simple case scenario is illustrated below.
+Markov chain Monte Carlo packege PYMC3 3.8\cite{pymc3} was used  with Python 3.7 for the simulation.
+Figures were generated by Python 3.7.
+The results were verified with JAGS 4.3.0 and R 3.6.2.
 
@@
 	\includegraphics[width=1\linewidth]{Figures/earlyresult1_ot.png}
-	\caption{Boxplot of posterior samples of $O_t$ (2000 samples for each month) with actual data points of simulated $O_t$ values. The simulated values are shown as red dots.}
+	\caption{Boxplot of posterior samples of $O_t$ (2000 samples for each month) with simulated $O_t$ values as red dots.}
 	\label{pst_ot}

### 2020-03-13 latex/newpaper.tex
histogram added latex

@@
 
+Figure \ref{hst_ot} is the histogram of posterior samples of total overdoses per year ($O_1 + O_2 + ...
++ O_{12}$).
+2000 samples were collected and shown as the histogram.
+It is seen that the plot is bell shaped and not skewed.
+The total number of overdoses per year is around 10900 to 11000.
+[mention about credible interval] \\
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{Figures/hist_sum_ot.png}
+	\caption{Histogram of posterior samples of total overdoses per year}
+	\label{hst_ot}
+\end{figure}
+
 \newpage

### 2020-04-01 Rcodes/JAGS_overdoses.Rmd
words style

@@
 ---
+# Make sure I used only useful posterior samples for the parameters.
+
 # Basic Model: Ambulance + Overdose
@@
 Now we suppose the population size of BC, N, is unknown.
-We may have some information about the population size and transform it into a prior distribution.
-Let's assume from the point that we don't have a good understanding of the sample size.
-But N should be at least bigger than the sample size of survey, the ambulance attended overdoses.
-We have 12 data points of $U_t$.
-We also have 12 data points of $x_A$.
+Notice that the prior distribution of N should reflect our prior knowledge about N which is even before we encounter the data set.
+Hence any information driven from the data set ($O_t, x_A$) cannot affect the prior of N.
+Suppose we have some information about the population size and transform it into a prior distribution.
 
-$$N \geq max(\vec{U_t}, \vec{x_A})$$
+$$N \sim N(10000,1000)$$
+The first prior we test is normal distribution where $\pm 20\%$ covers most of the support of N. Since $\pm2\sigma$ covers 95% of the distribution of Normal distribution, we set $\sigma = 1000$.
 
-Where $\vec{U_t}, \vec{x_A}$ are the vectors of $U_t$ and $x_A$.
+# Now the problem is that the trace plot of N.
 
-Then, we need a theoretical maximum value of N.
-BC is a province of Canada.
-For now, since I do not have a good idea about the maximum possible number of N, let's just say two times of the real value is the maximum.
+# todo: read the doing bayesian book and get the better idea about how to deal with this situation.
 
-$$N \in  \{a, a+1, a+2, ..., b\} \\ \textbf{where } a= \max(\vec{U_t}, n_A), b= 2 \times \textbf{real value of  N}$$
-
-However, in the model file below, we will use a prior such that 
-$$N \in  \{0,1,2, ..., b\} \\ \textbf{where }  b= 2 \times \textbf{real value of  N}$$ since we don't know how to construct such prior as described below.
-
+Notice that JAGS takes precison, which is $1/\sigma^2$ for its argument.
 ```{r Enter the models}
@@
 # Uncertain N takes in to the model
-N ~ dcat(p_i)
+N ~ dnorm(mu_N,(1/(sigma_N)^2))
 
@@
 ```{r Defining Values}
-p_i <- seq(1:(2*N))
 n_T <- length(df$o_t)
@@
 ```{r from variable name to \'variable_name\'}
-dat.N <- list(# prior for N
-            'p_i' = p_i,
+dat.N <- list(# priors for N
+            'mu_N' = 10000,
+            'sigma_N' = 1000,
+
             # priors for ambulance model
@@
 ```{r}
-test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 1000)
-interations = 1000
+test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 10000)
+interations = 10000
 burnin= floor(interations/2)
 summary(window(test), start=burnin)
-head(test)
+# head(test)
+```
+```{r}
+plot(test)
 ```
@@
 N <- pst_mtx.N[,1]
+length(N)
 ```
+
+```{r}
+plot(N )
+```
+
 ```{r}
 ggplot(data = as.data.frame(N), aes(x= N)) +geom_histogram()+ geom_vline(xintercept =10000)
-```
-
-It took a lot of time to fit the bayesian model this time.
-We have overestimated N value whose mode is around 17000 while the real value is 10000.
-Unif(0,20000) has 10000 as its mean.
-I guess the b was too large.
-I will wait for the reply from Paul. 
-```{r}
-plot(test)
 ```

### 2020-04-01 latex/newpaper.tex
words style

@@
-\documentclass[]{article}
+\PassOptionsToPackage{margin=1in}{geometry}
+\documentclass[12pt]{article}
+\usepackage{wordlike} 
+%\usepackage{mathptmx}
+%\usepackage[scaled=.90]{helvet}
+%\usepackage{courier}
 \usepackage{graphicx}
@@
 \graphicspath{{Figures/}} 
+\usepackage{setspace}
+
 \usepackage{hyperref}
@@
 \listoftables % Print the list of tables
+\doublespacing
 
@@
 This compares to 518 in 2015, an increase of 79.2\%.
-\cite{bccdc_opioid}  The goal of this project is to apply Bayesian evidence synthesis to better understand the opoid crisis in British Columbia, Canada.
+\cite{bccdc_opioid}  The goal of this project is to apply Bayesian evidence synthesis to better understand the opioid crisis in British Columbia, Canada.
 \\
@@
 One difficulty in coping with the opioid crisis is that the total number of overdoses is unknown.
-The reason is that the use of prescription opioids often leads to use of illicit opioids and the illicit usage cannot be known.
+The reason is that the use of prescription opioids often leads to use of illicit opioids and the extend of  illicit usage cannot be known.
 Many of those who become addicted to opioids do so after initially receiving a prescription.
@@
 This is often the time where people will turn to illicit drugs or other analogues.
-Because prescription opioids are so expensive, this is when many users turn to heroin.
+Because prescription opioids are so expensive,  many users turn to heroin.
 It is often cheaper, more potent, and easier to locate than what they were taking before.
@@
 Since the total number of overdoses is unknown, this number needs to be estimated.
-Bayesian statistics can be a way to approach the problem and give us a good estimate of the number [\textbf{defensible?
-depending on the data, a precise estimate may not be possible}].
+Bayesian statistics can be a way to approach the problem and give us a good estimate of the number with a defensible range of uncertainty.
 Bayesian statistics is a theory in the field of statistics based on the Bayesian interpretation of probability where probability expresses a degree of belief in an event.
-\cite{wiki_bayes} Detailed explanation is provided at the following section.
-[attempt: explain Bayesian statistics]\\
-
-All examples here were performed in Python 3.7 using the library pyMC (reference) and JAGS (reference).
-Training was performed using No U-Turn Sampling (NUTS) over two chains with 1000 iterations (is it sample size?).
-Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
-Data processing was carried out using the Pandas and SciPy library [reference].
-Data visualization was performed using the libraries Seaborn and Matplotlib [ref].
-Code for all examples in this study are provided.
-\\
+\cite{wiki_bayes} Further explanation is provided at the following section.
+[todo: explain Bayesian statistics a little bit more here]\\
 
@@
 \normalsize
-
+[\textbf{phrase differently: the prior and the data relate to all the unknown parameter, but as a last stage hou focuz on one target.}]
 The number of overdoses is our ultimate target of estimation.
-To achive the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
+To achieve the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
 The posterior belief is coming from both of these sources of the information.
@@
 \end{equation}
+[\textbf{Question: to make this flow better, maybe you want a version of (2) which generically describes parameter and latent varaibles given observed data.}]
 where $O_t$ is the total number of overdoses in a given month $t$ in British Columbia, and $y$ represents the collected data set providing relevant samples of the target variable.
@@
 The problem, however, is that the data $y$ is not available.
-Hence it is needed to approach the estimation in a indirect fashion.
+Hence it is needed to approach the estimation in an indirect fashion.
 Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
@@
 
+[\textbf{need more explicitness
+	t:month, considering T months all together}]
  The simplest conceptual model described in Figure \ref{simple_draw} is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
-\cite{Irvine:modelling} Then denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$.
-It is assumed that the total overdose $O_t$ follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
-Note that the population size, N, is unknown and vary over time in reality but it is assumed known constant for the investigator here for simplicity.
+\cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at time $t$.
+It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
+Note that the population size, N, is unknown and vary over time in reality but it is assumed known and constant for the investigator here for simplicity.
 
@@
 
-Since estimation of $O_t$ is not straightforward as none of the variables ($\mu$, $\sigma$, $N$) determining $O_t$  is known, we assume there was a survey conducted to infer $O_t$.
-The survey asked witnesses of opioid overdoses in danger whether someone called an ambulance.
-This provides estimate of the proportion of ambulance call $p_A$ among the overdoses.
-Let $n_{A}$ the sample size of the survey and $x_{A}$ to be the total number who confirmed they did call ambulance.
+Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is known, we assume there was a survey conducted to infer $O_t$.
+The survey [\textbf{survey doesn't do this directly, but it 'helps to support' inference about Ot}] asked witnesses of opioid overdoses whether someone called an ambulance.
+This provides an estimate of the proportion of ambulance call $p_A$ among the overdoses.
+For the moment $p_A$ is assumed constant across time for simplicity.
+Let $n_{A}$ be the sample size of the survey, and $x_{A}$ be the total number who confirmed that an ambulance was called.
 It is assumed that $x_{A}$ follows a Binomial distribution:
@@
 
-Let \(U_t\) the number of  ambulance-attended overdoses at a time point $t$.
+Let \(U_t\) be the number of  ambulance-attended overdoses at a time point $t$.
 In general, the data of ambulance-attended overdoses \(U_t\) can be obtained.
-It is assumed that  \(U_t\) follows Binomial distribution:
+It is assumed that  \(U_t\) follows a Binomial distribution:
 \begin{equation}
@@
 Note that now $O_t$ can be estimated because $p_A$ can be infered from the survey data and the data regarding $U_t$ is given.
-For the moment $p_A$ is assumed constant across time for simplicity.
 We suggest a simple model as a start where the model only combines Overdose Model (\ref{overdose}) and Ambulance Call-outs Model (\ref{ambulance}) with the relation (\ref{over_amb}).
@@
 \normalsize 
-The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.  \\
-
+The first simulation assumed $N= 10000, n_{A}=1000$.  \\
+[\textbf{Q;Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}]
 \subsubsection{Markov chain Monte Carlo}
@@
 \normalsize 
-There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets were simulated as follows. 
-It was assumed that the two data sets were collected for a year (t=1,2,3, ..., 12).
-The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months.
+There exist two data sets: the survey data ($n_A, x_A$), and the ambulance attended overdose data ($U_{1:T}$). The two data sets were simulated as follows. 
+It was assumed that the two data sets were collected for a year (T=12).
+The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months [one survey or 12?].
 The $x_A$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
-In terms of the overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
+In terms of the overdose data, it is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
 The vector of $O_t$ was generated following the overdose model (\ref{overdose}).
@@
 The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).
-Note that only $U_t$ and $x_A$ are known as the likelihood and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
+Note that only $U_t$ and $x_A$ are known as the likelihood [intended meaning?] and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 This leads the posterior distribution of variables of interest to heavily depend on the likelihood.
-Later, the noninformative priors will be changed and the impact of the changes over posteriors will be investigated.
+Later, the noninformative priors will be changed and the impact of the changes to posteriors will be investigated.
 \\
 
-\subsection{Early Result }
+\subsection{Initial Result}
+
+[\textbf{combine the two paragraph into one.}]
+All examples here were performed in Python 3.7 using the library pyMC (reference) and JAGS (reference).
+Training was performed using No U-Turn Sampling (NUTS) over two chains with 1000 iterations (is it sample size?).
+Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
+Data processing was carried out using the Pandas and SciPy library [reference].
+Data visualization was performed using the libraries Seaborn and Matplotlib [ref].
+Code for all examples in this study are provided.
 

### 2020-04-04 latex/newpaper.tex
edited the major parts of latex

@@
 \normalsize
-[\textbf{phrase differently: the prior and the data relate to all the unknown parameter, but as a last stage hou focuz on one target.}]
+[\textbf{phrase differently: the prior and the data relate to all the unknown parameter, but as a last stage you focus on one target.}]
 The number of overdoses is our ultimate target of estimation.
@@
 \end{equation}
-[\textbf{Question: to make this flow better, maybe you want a version of (2) which generically describes parameter and latent varaibles given observed data.}]
+[\textbf{Question: to make this flow better, maybe you want a graph figure version of (1) which generically describes parameter and latent varaibles given observed data.}]
 where $O_t$ is the total number of overdoses in a given month $t$ in British Columbia, and $y$ represents the collected data set providing relevant samples of the target variable.
@@
 Hence it is needed to approach the estimation in an indirect fashion.
-Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
 \\
@@
 
-[\textbf{need more explicitness
+[\textbf{attempted: need more explicitness
 	t:month, considering T months all together}]
- The simplest conceptual model described in Figure \ref{simple_draw} is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
-\cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at time $t$.
+Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
+It is assumed that we are interested in the opioid crisis from month 1 to month T; subscript 1:T for any variable denotes the variable of every target months and subscription t denotes month t.
+The simplest conceptual model described in Figure \ref{simple_draw} is to take an underlying log-rate $z_t$ where that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
+\cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at month $t$.
 It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
@@
 
-Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is known, we assume there was a survey conducted to infer $O_t$.
-The survey [\textbf{survey doesn't do this directly, but it 'helps to support' inference about Ot}] asked witnesses of opioid overdoses whether someone called an ambulance.
+Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is not known, we assume there was a survey conducted to help to support inference about  $O_t$.
+The survey asked witnesses of opioid overdoses whether someone called an ambulance.
 This provides an estimate of the proportion of ambulance call $p_A$ among the overdoses.
@@
 \normalsize 
-The first simulation assumed $N= 10000, n_{A}=1000$.  \\
-[\textbf{Q;Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}]
+The first simulation assumed $N= 10000, n_{A}=1000$.  [\textbf{Q;Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}]\\
+
 \subsubsection{Markov chain Monte Carlo}
@@
 \normalsize 
-There exist two data sets: the survey data ($n_A, x_A$), and the ambulance attended overdose data ($U_{1:T}$). The two data sets were simulated as follows. 
+There exist two data sets: the survey data ($n_A, x_A$), and the ambulance attended overdose data ($U_{1:T}$).
+The two data sets were simulated as follows.
 It was assumed that the two data sets were collected for a year (T=12).
-The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months [one survey or 12?].
+In terms of the survey data, it is assumed that  we have 12 survey data sets where each survey data set was collected for each month .
+The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months.
 The $x_A$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
@@
 The vector of $U_t$ was generated from the Binomial relation of the two variables ($\ref{over_amb}$).
-The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).
-Note that only $U_t$ and $x_A$ are known as the likelihood [intended meaning?] and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
+The two generated vectors have the same length with the survey data (T=12).
+Note that only $U_t$ and $x_A$ are the only known as the data set in real world and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 
-Figure \ref{hst_ot} is the histogram of posterior samples of total overdoses per year ($O_1 + O_2 + ...
+Let $O_T$ denote the total overdose per year ($O_T = O_1 + O_2 + O_3 + ...
 + O_{12}$).
+Figure \ref{hst_ot} is the histogram of posterior samples of total overdoses $O_T$ per year.
 2000 samples were collected and shown as the histogram.
 It is seen that the plot is bell shaped and not skewed.
-The total number of overdoses per year is around 10900 to 11000.
-[mention about credible interval] \\
+The total number of overdoses per year is inferred to be around 10900 to 11000.
+[mention about credible interval] [\textbf{not on density scale (count only has meaning relative to chosen MCMC number iterations))}]\\
 
@@
 \subsubsection{Posterior Predictive Check}
-Posterior predictive checking is a model validation technique that we simulate some replicated data under the fitted model then compare the new data to the observed data.
+Posterior predictive checking is a model validation technique when we simulate some replicated data under the fitted model then compare the new data to the observed data.
 If the model fits, then replicated data generated under the model should look similar to observed data.
@@
 Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$.
-It is shown that the posterior predictive estimates of $U_t$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $O_t$.
+It is shown that the posterior predictive estimates of $U_t$ are failry accurate, for the same two reasons regarding the accuracy of the posterior distribution of $O_t$.
 It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11).
-Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; for months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\
+Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; for months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11).\\
 
@@
 	\subfloat[ppc of $p_A$]{\includegraphics[width=.50\columnwidth]{early_r_ppc1_xt}\label{ppc_xt}}
-	\caption[two early result box plots:]{Boxplots of posterior predictive samples of $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value.
+	\caption[two early result box plots:]{Boxplots of posterior predictive samples of $U_t$ and $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value.
 The simulated values are shown as a red dots.}
@@
 Figure \ref{ppc_xt} is the boxplots of posterior predictive samples of $x_A$.
-It is shown that the posterior predictive estimates of $x_A$ is tolerably accurate since (1) the boxplots contain actual values of $O_t$ within their lines connecting the maximum and the minimum and (2) the ranges of IQR seem modereately narrow.
+It is shown that the posterior predictive estimates of $x_A$ are tolerably accurate since (1), the boxplots contain actual values of $O_t$ within their lines connecting the maximum and the minimum, and (2), the ranges of IQR seem narrow.
 \\
 
-\subsection{Early Result: Contamination of $p_A$ } 
-One of the attention of this research project is to investigate how robust the model is from contaminations of the data sets.
+\subsection{Initial Result: Contamination of $p_A$ } 
+One focus of this research project is to investigate how robust the model is to contaminations of the data sets.
 The first inspection is to check an impact of a contamination of $p_A$; what would happen if the estimation of $p_A$ is biased?
 It is assumed that the survey data gives us a wrong estimate of $p_A$ such that it would be underestimated or overestimated.
-We then want to see how the biased estimation of $p_A$ affects the estimate of $O_t$, the total overdose.\\
+We then want to see how the biased estimation of $p_A$ affects the estimate of $O_t$, the total number of overdoses.\\
 
-Both of underestimation and overestimation were conducted for the analysis.
-In terms of underestimation, the simulated survey data ($n_A, x_A$) was generated with $p_A=0.6$ while the true value of $p_A$ is 0.8, and all the other assumptions hold the same.
+Both underestimation and overestimation were considered for the analysis.
+In terms of underestimation, the simulated survey data ($n_A, x_A$) was generated with $p_A=0.6$ while the true value of $p_A$ is 0.8, and all the other assumptions hold as before.
 That is, $x_A$ is generated from $x_A \sim Bin(n_A, 0.6)$, while $U_t$ is generated from $U_t \sim Bin(O_t, 0.8)$ for every $t$.
@@
 It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots.
-This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); given the likelihood of the ambulance attended overdoses, $U_t$, a value of $O_t$ is an unknown parameter of the binomial distribution, $Bin(O_t, p_A)$.
-Let $o_t$, $u_t$ are samples from $O_t$ and the likelihood value of $U_t$ respectively.
-Then $o_t$ is proportional to the quantity generated by multiplying $u_t$ and the inverse of $\hat{p}_{A}$ where $\hat{p}_{A}$ is underestimated estimator of $p_A$; hence $\frac{1}{\hat{p}_{A}}$ is overestimated which leads overestimation of $O_t$.
+This is explainable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); given the likelihood of the ambulance attended overdoses, $U_t$, a value of $O_t$ is an unknown parameter of the binomial distribution, $Bin(O_t, p_A)$.
+Let $o_t$, $u_t$ are posterior samples from $O_t$ and the observed value of $U_t$ respectively.
+Then $o_t$ is proportional to the quantity generated by multiplying $u_t$ and the inverse of $\hat{p}_{A}$ where $\hat{p}_{A}$ is underestimated estimator of $p_A$; hence $\frac{1}{\hat{p}_{A}}$ is overestimated which leads overestimation of $O_t$:
 \begin{equation}
@@
 
-Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ leads an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
+Figure \ref{contam_ut} gives the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
+It is seen that none of the contaminations of $p_A$ leads to a an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
 
@@
 The result also matches with our intution.
-Posterior predictive checkes use the existing data points twice; firstly it uses the data to obtain the posterior samples of paramaters, and secondly the samples are used to produce posterior predictive samples.
+Posterior predictive checks use the existing data points twice; firstly it uses the data to obtain the posterior samples of paramaters, and secondly the samples are used to produce posterior predictive samples.
 Given the fact that there could be an overfitting issue due to using the data twice, the result shown here seems reasonable.
@@
 
-Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
+Figure \ref{contam_xt}  gives the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
 It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.

### 2020-04-04 latex/newpaper.tex
latex double checked before asking questions

@@
 One difficulty in coping with the opioid crisis is that the total number of overdoses is unknown.
-The reason is that the use of prescription opioids often leads to use of illicit opioids and the extend of  illicit usage cannot be known.
+The reason is that the use of prescription opioids often leads to use of illicit opioids and the extent of illicit usage cannot be known.
 Many of those who become addicted to opioids do so after initially receiving a prescription.
@@
 \normalsize
-[\textbf{phrase differently: the prior and the data relate to all the unknown parameter, but as a last stage you focus on one target.}]
+[\textbf{phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]
 The number of overdoses is our ultimate target of estimation.
@@
 
-[\textbf{attempted: need more explicitness
-	t:month, considering T months all together}]
 Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
@@
 
-Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is not known, we assume there was a survey conducted to help to support inference about  $O_t$.
-The survey asked witnesses of opioid overdoses whether someone called an ambulance.
-This provides an estimate of the proportion of ambulance call $p_A$ among the overdoses.
-For the moment $p_A$ is assumed constant across time for simplicity.
-Let $n_{A}$ be the sample size of the survey, and $x_{A}$ be the total number who confirmed that an ambulance was called.
+Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is not known, we assume there was a survey conducted to help to support the inference about  $O_t$.
+The survey asked witnesses of opioid overdoses whether someone called an ambulance for each month.
+This provides an estimate of the proportion of ambulance call $p_{A,t}$ among the overdoses.
+For the moment $p_{A,t}$ is assumed constant across time for simplicity.
+Let $n_{A,t}$ be the sample size of the survey, and $x_{A,t}$ be the total number who confirmed that an ambulance was called.
 It is assumed that $x_{A}$ follows a Binomial distribution:
@@
 \left.\begin{aligned}
-x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
+x_{A,t} \sim Bin(n_{A,t},p_{A,t})\end{aligned}\right.
 \text{ambulance call-outs model}
@@
 \left.
-U_t \sim Bin(O_t, p_A)
+U_t \sim Bin(O_t, p_{A,t})
 \right.
@@
 
-Note that now $O_t$ can be estimated because $p_A$ can be infered from the survey data and the data regarding $U_t$ is given.
+Note that now $O_t$ can be estimated because $p_{A,t}$ can be infered from the survey data and the data regarding $U_t$ is given.
 We suggest a simple model as a start where the model only combines Overdose Model (\ref{overdose}) and Ambulance Call-outs Model (\ref{ambulance}) with the relation (\ref{over_amb}).
@@
 \normalsize 
-There exist two data sets: the survey data ($n_A, x_A$), and the ambulance attended overdose data ($U_{1:T}$).
+There exist two data sets: the survey data ($n_{A,1:T}, x_{A,1:T}$), and the ambulance attended overdose data ($U_{1:T}$).
 The two data sets were simulated as follows.
@@
 In terms of the survey data, it is assumed that  we have 12 survey data sets where each survey data set was collected for each month .
-The true value of $p_A$ was set as $p_A=0.8$ and the survey sample size $n_A$ was set as $n_A=1000$ for every months.
-The $x_A$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
+The true value of $p_{A,t}$ was set as $p_{A,t}=0.8$ and the survey sample size $n_{A,t}$ was set as $n_{A,t}=1000$ for every months (t=1:T).
+The $x_{A,t}$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
 In terms of the overdose data, it is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
@@
 The two generated vectors have the same length with the survey data (T=12).
-Note that only $U_t$ and $x_A$ are the only known as the data set in real world and $p_A$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
+Note that only $U_t$ and $x_{A,t}$ are the only known as the data set in real world and $p_{A,t}$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 
-Let $O_T$ denote the total overdose per year ($O_T = O_1 + O_2 + O_3 + ...
-+ O_{12}$).
-Figure \ref{hst_ot} is the histogram of posterior samples of total overdoses $O_T$ per year.
+Let $O_T$ denote the total overdose per year ($O_T = \sum_{t=1}^{12}O_t $).
+Figure \ref{hst_ot} is the histogram of posterior samples of the total number of overdoses $O_T$ per year.
 2000 samples were collected and shown as the histogram.
@@
 This is explainable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); given the likelihood of the ambulance attended overdoses, $U_t$, a value of $O_t$ is an unknown parameter of the binomial distribution, $Bin(O_t, p_A)$.
-Let $o_t$, $u_t$ are posterior samples from $O_t$ and the observed value of $U_t$ respectively.
+Let $o_t$, $u_t$ denote posterior samples from $O_t$ and the observed value of $U_t$ respectively.
 Then $o_t$ is proportional to the quantity generated by multiplying $u_t$ and the inverse of $\hat{p}_{A}$ where $\hat{p}_{A}$ is underestimated estimator of $p_A$; hence $\frac{1}{\hat{p}_{A}}$ is overestimated which leads overestimation of $O_t$:
@@
 Figure \ref{contam_ut} gives the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ leads to a an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
+It is seen that none of the contaminations of $p_A$ leads to an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
 

### 2020-04-08 latex/newpaper.tex
edit on the main part rephrasing

@@
 \usepackage{subfig}
+\usepackage{soul}
 \graphicspath{{Figures/}} 
@@
 \normalsize
-[\textbf{phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]
+[\textbf{phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]\\
+
+[\textbf{following the comment, the first attempt:}]\\  
 The number of overdoses is our ultimate target of estimation.
@@
 The posterior belief is coming from both of these sources of the information.
+Let $\theta$ denote all the unknown variables and let $Y$ denote all the related data sets.
+The following equation express the idea of Bayesian generative models in a mathmatical form:
+\begin{equation}
+\label{generic_bayes_thm}
+\left.\begin{aligned}
+p(\theta| Y ) = \frac{ p(\theta, Y)   }{ p(y)} = \frac{ p(\theta) p(Y|\theta)}{ p(Y)} \propto p(\theta) p(Y| \theta) \end{aligned}\right.
+\end{equation}
+[\textbf{Question: to make this flow better, maybe you want a  version of (1) which generically describes parameter and latent varaibles given observed data.}]
+The equation \ref{generic_bayes_thm} states that the posterior belief of unobserved variables, $p(\theta|Y)$, is proportional to the mixture of the prior belief of the varaibles, $p(\theta)$, and the related data sets given the prior belief, $p(Y|\theta)$.
+Notice that the vector $\theta$ from equation \ref{generic_bayes_thm} includes parameters and latent variables, and hence it contains the total number of overdoses.
+
+Let us assume that we are interested in the opioid crisis from month 1 to month T.
+Let the subscript 1:T for any variable denote the variable of every target months and subscription t for any variable denote the variable of the month t.
+Then the target variable, the total number of overdoses from month 1 to month T  in British Columbia, is denoted by $O_{1:T}$.
+\iffalse
+Let $O_t$ denote the total number of overdoses in a given month $t$. 
+\fi
+In Bayesian concepts, the variable of interest $O_{1:T}$ is a vector of fixed number but the numbers are yielded from a certain distribution where the distribution describes knowledge about $O_{1:T}$; that is, each $O_t$ is considered to be a sample of a random variable.
+The simplest conceptual model is to take an underlying log-rate $z_t$ for each month $t$ where $z_t$ is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
+\cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at month $t$.
+It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
+Note that the population size, N, is unknown and vary over time in reality but it is assumed known and constant for the investigator here for simplicity.
+The mechanism is named as overdose model  and is illustrated in mathmatical form below:
+
+\iffalse
+{The number of overdoses is our ultimate target of estimation.
+To achieve the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
+The posterior belief is coming from both of these sources of the information.
 The following equation express the idea in a mathmatical form:
+\fi
+
+\iffalse
 \begin{equation}
@@
 \end{equation}
-[\textbf{Question: to make this flow better, maybe you want a graph figure version of (1) which generically describes parameter and latent varaibles given observed data.}]
+\fi
+
+\iffalse
 where $O_t$ is the total number of overdoses in a given month $t$ in British Columbia, and $y$ represents the collected data set providing relevant samples of the target variable.
@@
 Hence it is needed to approach the estimation in an indirect fashion.
-\\
-\begin{figure}[htb]
-	\centering
-	\includegraphics[width=0.5\linewidth]{Figures/simple_drawing}
-	\caption{Example of estimating the total overdoses indirectly given two data sets: ambulance attended overdoses and survey data.}
-	\label{simple_draw}
-\end{figure}
+\fi
 
-Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
+\iffalse
  It is assumed that we are interested in the opioid crisis from month 1 to month T; subscript 1:T for any variable denotes the variable of every target months and subscription t denotes month t.
@@
 Note that the population size, N, is unknown and vary over time in reality but it is assumed known and constant for the investigator here for simplicity.
+\fi
 
@@
 For the moment $p_{A,t}$ is assumed constant across time for simplicity.
-Let $n_{A,t}$ be the sample size of the survey, and $x_{A,t}$ be the total number who confirmed that an ambulance was called.
-It is assumed that $x_{A}$ follows a Binomial distribution:
+Let $n_{A,t}$ be the sample size of the survey at month $t$, and $x_{A,t}$ be the total number who confirmed that an ambulance was called at the month $t$.
+It is assumed that $x_{A,t}$ follows a Binomial distribution with number of trials, $n_{A,t}$, and  success probability for each trial, $p_{A,t}$.
+The mechanism is named as ambulance call-outs model and is illustrated in mathmatical form below:
 
@@
 
+Now the estimated $p_{A,t}$ can help to infer $O_t$ if there is a data regarding the number of  ambulance-attended overdoses.
 Let \(U_t\) be the number of ambulance-attended overdoses at a time point $t$.
@@
 
-Note that now $O_t$ can be estimated because $p_{A,t}$ can be infered from the survey data and the data regarding $U_t$ is given.
-We suggest a simple model as a start where the model only combines Overdose Model (\ref{overdose}) and Ambulance Call-outs Model (\ref{ambulance}) with the relation (\ref{over_amb}).
+Note that $O_t$ can be estimated because $p_{A,t}$ can be infered from the survey data and the data regarding $U_t$ is given.
+Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
+The variables that can be observed (N, $U_{1:T}$, and $X_{A,1:T}$) are drawn in the square box and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, P_{A,1:T})$) that are not observarble from possible data sets are drawn in the circle box.
+We suggest a simple model as a start where the model only combines Overdose Model \ref{overdose} and Ambulance Call-outs Model \ref{ambulance} with the relation \ref{over_amb}.
 The next step is to run some simulations to figure out how different types of inputs lead some changes of output.
-\\
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=200mm,scale=0.5]{Figures/simple_drawing}
+	\caption{Example of estimating the total number of overdoses indirectly given two data sets: ambulance attended overdoses and survey data.}
+	\label{simple_draw}
+\end{figure}
 

### 2020-04-08 latex/newpaper.tex
ready for tomorrow; did what I understood

@@
 Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
-The variables that can be observed (N, $U_{1:T}$, and $X_{A,1:T}$) are drawn in the square box and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, P_{A,1:T})$) that are not observarble from possible data sets are drawn in the circle box.
-We suggest a simple model as a start where the model only combines Overdose Model \ref{overdose} and Ambulance Call-outs Model \ref{ambulance} with the relation \ref{over_amb}.
-The next step is to run some simulations to figure out how different types of inputs lead some changes of output.
+The variables that can be observed (N, $U_{1:T}$, and $x_{A,1:T}$) are drawn in the square boxes and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, p_{A,1:T} $) that are not observarble from possible data sets are drawn in the circle boxes.
+Variables that need prior distributions have asterisks on their draws.
+The arrows indicate which variables affects which variables; for example, there is an arrow from $\mu, \sigma^2$ to $\lambda_{1:T}$ since $\mu, \sigma^2$ is the parameters of $\lambda_{1:T}$, and the two arrows from  $\lambda_{1:T}$ and $N$ to $O_{1:T}$ indicate that both $N$ and $\lambda_{1:T}$ affects the value of $O_{1:T}$ since the product of $\lambda_{1:T}$ and $N$ is the parameter of the distribution of $O_{1:T}$.
 
@@
 \end{figure}
+
+We suggest a simple model in figure \ref{simple_draw} as a start where the model only combines Overdose Model \ref{overdose} and Ambulance Call-outs Model \ref{ambulance} with the relation \ref{over_amb}.
+The next step is to run some simulations to figure out how different types of inputs lead some changes of output.
 

### 2020-04-08 latex/newpaper.tex
generative model written

@@
 [\textbf{Question: to make this flow better, maybe you want a  version of (1) which generically describes parameter and latent varaibles given observed data.}]
+where $p(\cdot)$ is probability density (or mass) function.
 The equation \ref{generic_bayes_thm} states that the posterior belief of unobserved variables, $p(\theta|Y)$, is proportional to the mixture of the prior belief of the varaibles, $p(\theta)$, and the related data sets given the prior belief, $p(Y|\theta)$.
@@
 The next step is to run some simulations to figure out how different types of inputs lead some changes of output.
+
+\subsection{Generative Models}
+
+[\textbf{draft:}] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
+
+\begin{equation}
+\label{gen_model}
+\left.\begin{aligned}
+p(\theta) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{A,t}, n_{A,t}, x_{A,t}, U_t) \\
+& = p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{A,t})p(n_{A,t})p(x_{A,t}|p_{A,t},n_{A,t})p(U_t|O_t, p_{A,t}) \\
+& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(O_t|\lambda_t, N)p(p_{A,t})p(x_{A,t}|p_{A,t},n_{A,t})p(U_t|O_t, p_{A,t}) 
+ \end{aligned}\right.
+\end{equation}
 

### 2020-04-12 latex/researchpaper.tex
latex editing ongoing before simulation

--- 2020-04-08 latex/newpaper.tex
@@
 \normalsize
-[\textbf{phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]\\
+[\textbf{I guess its done; phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]\\
 
-[\textbf{following the comment, the first attempt:}]\\  
+[\textbf{done?; following the comment, the first attempt:}]\\  
 The number of overdoses is our ultimate target of estimation.
@@
 \end{equation}
-[\textbf{Question: to make this flow better, maybe you want a  version of (1) which generically describes parameter and latent varaibles given observed data.}]
+[\textbf{done?: to make this flow better, maybe you want a  version of (1) which generically describes parameter and latent varaibles given observed data.}]
 where $p(\cdot)$ is probability density (or mass) function.
@@
 
-[\textbf{draft:}] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
+[\textbf{done?: mentioning how generative model factors and mentioning fig 1 again.}] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
 
@@
 \left.\begin{aligned}
-p(\theta) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{A,t}, n_{A,t}, x_{A,t}, U_t) \\
-& = p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{A,t})p(n_{A,t})p(x_{A,t}|p_{A,t},n_{A,t})p(U_t|O_t, p_{A,t}) \\
-& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(O_t|\lambda_t, N)p(p_{A,t})p(x_{A,t}|p_{A,t},n_{A,t})p(U_t|O_t, p_{A,t}) 
+p(\theta) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{(A,t)}, n_{(A,t)}, x_{(A,t)}, U_t) \\
+& = p(\mu)p(\sigma^2|\mu)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{(A,t)})p(n_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) \\
+& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(O_t|\lambda_t, N)p(p_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) 
  \end{aligned}\right.
 \end{equation}
+
+The chain rule for random variables \cite{chain_rule} is used to the joint distribution of all variables at the first line of equation \ref{gen_model} such that the second line of the equation is derived.
+A priori independence assumptions and conditional independence assumptions were presumed so that the second line of the equation is simplified.
+For example, it is asumed that  $\sigma^2$ and $\mu$ are independent to each other and $p_{A,t}$ and also all the variables from overdose model \ref{overdose}, ($\mu, \sigma^2, \lambda_t, N, O_t$), are independent to each other.
+Note that as a function of unknown variables, the function from the second line of the equation \ref{gen_model} is proportional to the function from the third line of the equation since $N$ and $n_{(A,t)}$ are presumed to be known and constants such that the density functions of the known variables are merely constants.
+Therefore the third line of the equation is the simplified generative model of interest.
 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{alpha}
 \bibliography{sample.bib}

### 2020-04-13 Rcodes/JAGS_overdoses.Rmd
on my way to Monte Carlo

@@
 df$X <- NULL 
-head(df)
+# head(df)
+
+knitr::kable(df, "pandoc")
 ```

### 2020-04-13 latex/researchpaper.tex
on my way to Monte Carlo

@@
 
-[\textbf{done?: mentioning how generative model factors and mentioning fig 1 again.}] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
+[\textbf{done?: comments: Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}] The amalgamation of the prior distribution and the statistical model can be referred to as the generative model.[reference] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
 
@@
 \left.\begin{aligned}
-p(\theta) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{(A,t)}, n_{(A,t)}, x_{(A,t)}, U_t) \\
+p(\theta, Y) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{(A,t)}, n_{(A,t)}, x_{(A,t)}, U_t) \\
 & = p(\mu)p(\sigma^2|\mu)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{(A,t)})p(n_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) \\
@@
 
+where $\theta = (\mu, ,\sigma^2,\lambda_t, O_t, p_{(A,t)})$ and $ Y = ( N, n_{(A,t)}, x_{(A,t)}, U_t)$.
 The chain rule for random variables \cite{chain_rule} is used to the joint distribution of all variables at the first line of equation \ref{gen_model} such that the second line of the equation is derived.
@@
 Note that as a function of unknown variables, the function from the second line of the equation \ref{gen_model} is proportional to the function from the third line of the equation since $N$ and $n_{(A,t)}$ are presumed to be known and constants such that the density functions of the known variables are merely constants.
-Therefore the third line of the equation is the simplified generative model of interest.
+Therefore the third line of the equation is the ultimate generative model of interest.
 
@@
 \normalsize 
-The first simulation assumed $N= 10000, n_{A}=1000$.  [\textbf{Q;Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}]\\
-
-\subsubsection{Markov chain Monte Carlo}
-The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
-The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived.
-\textbf{[Further explanation is going to be added.]} [attempt:brief explaination of MCMC and related topics]\\
+Simulation can be used to predict the performance of the Bayesian models.
+In order to measure the performance of the simplified model, a data set was simulated and the process is described in this chapter.
 
@@
 The $x_{A,t}$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
-In terms of the overdose data, it is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
+In terms of the overdose data, it is assumed that the total number of population were $N = 10000$ the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
 The vector of $O_t$ was generated following the overdose model (\ref{overdose}).
@@
 Later, the noninformative priors will be changed and the impact of the changes to posteriors will be investigated.
-\\
+
+\subsubsection{Markov Chain Monte Carlo}
+
+[Why do we need MCMC? Because we don't have platinum, golden, and silver grails.]\\
+Posterior distribution of the unseen variables, parameters ($\mu, \sigma^2, p_{A,1:T}$) and the latent variables ($\lambda_{1:T}, O_{1:T}$), given the seen variables, $Y$=($N, n_{(A,1:T)} x_{(A,1:T)}, U_{1:T}$), is the key to make any inference for any of the unseen variables, $\theta$.
+For the inference about the variable of interest, $O_{1:T}$, the marginal posterior distribution, $p(O_{1:T}|Y)$, is essential which can be obtained by integrating the joint posterior distribution, $p(\theta|Y)$, over all the other unseen variables except $O_{1:T}$.
+However, in general it is severely complicated and almost impossible to perform such multiple integrations because $p(\theta|Y)$ is a complex distribution of high dimensions.
+Moreover, $p(O_{1:T}|Y)$ may not have closed form nor follow a known parametric distribution, which lead the general inference procedure laborious.\\
+Markov Chain Monte Carlo (MCMC) is a detour to obtain posterior samples from $p(\theta|Y)$ without explicitly knowing the distribution.
+MCMC technique allows us to generate Monte Carlo realizations of samples, $\theta_{(i)}$, which are dependent, with the distribution of each sample, $p(\theta_{(i)})$, converging to $p(\theta|Y)$.\cite{paul}.\\
+[general explanation of what is MC and MC, how it works.]
+
+ Generative model auto pilot conceptunally there's only\\
+ one distribution of unseen stuff given the seen stuff we can use the distribution to make whatever inference claims that we need to make, such as point estimation, credibal interval, inference about parameters or latent variables.
+
+[Two types of bronze grails, NUTS, metropolis]
+
+[explain all the figures below was derived by NUTS, and python]
+
+The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
+The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived.
+\textbf{[Further explanation is going to be added.]} [attempt:brief explaination of MCMC and related topics]
 
@@
 
-Let $O_T$ denote the total overdose per year ($O_T = \sum_{t=1}^{12}O_t $).
-Figure \ref{hst_ot} is the histogram of posterior samples of the total number of overdoses $O_T$ per year.
+Let $O_+$ denote the total overdose per year ($O_+ = \sum_{t=1}^{12}O_t $).
+Figure \ref{hst_ot} is the histogram of posterior samples of the total number of overdoses $O_+$ per year.
 2000 samples were collected and shown as the histogram.
@@
 The total number of overdoses per year is inferred to be around 10900 to 11000.
-[mention about credible interval] [\textbf{not on density scale (count only has meaning relative to chosen MCMC number iterations))}]\\
+[mention about credible interval] \\
 

### 2020-04-23 latex/researchpaper.tex
comments applied on the plane

@@
 %opening
-\title{Bayesian Evidence Synthesis:opioid crisis}
-\author{Hyeongcheol Park* \& Paul Gustafson* \& Micheal A Irvine*\textsuperscript{1}}
+\title{Bayesian Evidence Synthesis:Opioid Crisis}
+\author{Hyeongcheol Park*}
 
@@
 \normalsize 
-[\textbf{I guess its done; phrase differently: the prior and the data relate to "all the unknown parameter", but as a last stage you focus on one target.}]\\
-
-[\textbf{done?; following the comment, the first attempt:}]\\  
 The number of overdoses is our ultimate target of estimation.
@@
 \left.\begin{aligned}
-p(\theta| Y ) = \frac{ p(\theta, Y)   }{ p(y)} = \frac{ p(\theta) p(Y|\theta)}{ p(Y)} \propto p(\theta) p(Y| \theta) \end{aligned}\right.
+p(\theta| Y ) = \frac{ p(\theta, Y)   }{ p(Y)} = \frac{ p(\theta) p(Y|\theta)}{ p(Y)} \propto p(\theta) p(Y| \theta), \end{aligned}\right.
 \end{equation}
-[\textbf{done?: to make this flow better, maybe you want a  version of (1) which generically describes parameter and latent varaibles given observed data.}]
+
 where $p(\cdot)$ is probability density (or mass) function.
-The equation \ref{generic_bayes_thm} states that the posterior belief of unobserved variables, $p(\theta|Y)$, is proportional to the mixture of the prior belief of the varaibles, $p(\theta)$, and the related data sets given the prior belief, $p(Y|\theta)$.
+The equation \ref{generic_bayes_thm} states that the posterior belief of unobserved variables, $p(\theta|Y)$, is proportional to the mixture of the prior belief of the variables, $p(\theta)$, and the related data sets given the prior belief, $p(Y|\theta)$.
 Notice that the vector $\theta$ from equation \ref{generic_bayes_thm} includes parameters and latent variables, and hence it contains the total number of overdoses.
@@
 Let us assume that we are interested in the opioid crisis from month 1 to month T.
-Let the subscript 1:T for any variable denote the variable of every target months and subscription t for any variable denote the variable of the month t.
+Let the subscript 1:T for any variable denote the variable for all target months and subscript t for any variable denote the variable of the month t.
 Then the target variable, the total number of overdoses from month 1 to month T  in British Columbia, is denoted by $O_{1:T}$.
@@
 \fi
-In Bayesian concepts, the variable of interest $O_{1:T}$ is a vector of fixed number but the numbers are yielded from a certain distribution where the distribution describes knowledge about $O_{1:T}$; that is, each $O_t$ is considered to be a sample of a random variable.
+In Bayesian concepts, the variable of interest $O_{1:T}$ is a vector of fixed numbers but the numbers are yielded from a certain distribution where the distribution describes knowledge about $O_{1:T}$; that is, each $O_t$ is considered to be a sample of a random variable.
 The simplest conceptual model is to take an underlying log-rate $z_t$ for each month $t$ where $z_t$ is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
 \cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at month $t$.
-It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the population of the region of interest is $N$.
+It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the size of the population in the region of interest is $N$.
 Note that the population size, N, is unknown and vary over time in reality but it is assumed known and constant for the investigator here for simplicity.
-The mechanism is named as overdose model  and is illustrated in mathmatical form below:
+The mechanism is named as the overdose model  and is illustrated in mathmatical form below:
 
@@
  \left.\begin{aligned}
- z_{t} \sim N(\mu, \sigma^{2}) \\
- \lambda_{t}^{OD} = \exp(z_{t})\\
+ z_{t} \sim N(\mu, \sigma^{2}), \\
+ \lambda_{t}^{OD} = \exp(z_{t}),\\
  O_{t} \sim Poi(\lambda_{t}^{OD}N)
  \end{aligned}\right\} 
- \text{		overdose model} 
+ \text{(overdose model).} 
  \end{equation}
@@
 Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is not known, we assume there was a survey conducted to help to support the inference about  $O_t$.
-The survey asked witnesses of opioid overdoses whether someone called an ambulance for each month.
+The survey asked witnesses of opioid overdoses whether someone called an ambulance, for each month.
 This provides an estimate of the proportion of ambulance call $p_{A,t}$ among the overdoses.
@@
 It is assumed that $x_{A,t}$ follows a Binomial distribution with number of trials, $n_{A,t}$, and  success probability for each trial, $p_{A,t}$.
-The mechanism is named as ambulance call-outs model and is illustrated in mathmatical form below:
+The mechanism is named as the ambulance call-outs model and is illustrated in mathmatical form below:
 
@@
 x_{A,t} \sim Bin(n_{A,t},p_{A,t})\end{aligned}\right.
-\text{ambulance call-outs model}
+\text{(ambulance call-outs model).}
 \end{equation}
@@
 \left.
-U_t \sim Bin(O_t, p_{A,t})
+U_t \sim Bin(O_t, p_{A,t}).
 \right.
@@
 Note that $O_t$ can be estimated because $p_{A,t}$ can be infered from the survey data and the data regarding $U_t$ is given.
-Figure \ref{simple_draw} shows an example of how the estimation can be proceeded.
-The variables that can be observed (N, $U_{1:T}$, and $x_{A,1:T}$) are drawn in the square boxes and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, p_{A,1:T} $) that are not observarble from possible data sets are drawn in the circle boxes.
-Variables that need prior distributions have asterisks on their draws.
-The arrows indicate which variables affects which variables; for example, there is an arrow from $\mu, \sigma^2$ to $\lambda_{1:T}$ since $\mu, \sigma^2$ is the parameters of $\lambda_{1:T}$, and the two arrows from  $\lambda_{1:T}$ and $N$ to $O_{1:T}$ indicate that both $N$ and $\lambda_{1:T}$ affects the value of $O_{1:T}$ since the product of $\lambda_{1:T}$ and $N$ is the parameter of the distribution of $O_{1:T}$.
+Figure \ref{simple_draw} shows an example of how the estimation can proceed.
+The variables that can be observed (N, $U_{1:T}$, and $x_{A,1:T}$) are drawn in squares and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, p_{A,1:T} $) that are not observarble from possible data sets are drawn in circles.
+Variables that need prior distributions are indicated with asterisks.
+The arrows indicate which variables affects which variables; for example, there is an arrow from $(\mu, \sigma^2)$ to $\lambda_{1:T}$ since $\mu, \sigma^2$ are the parameters of $\lambda_{1:T}$, and the two arrows from  $\lambda_{1:T}$ and $N$ to $O_{1:T}$ indicate that both $N$ and $\lambda_{1:T}$ affects the value of $O_{1:T}$ since the product of $\lambda_{1:T}$ and $N$ is the parameter of the distribution of $O_{1:T}$.
 
@@
 \subsection{Generative Models}
-
-[\textbf{done?: comments: Before moving on to simulation, you should write out how the generative(generable) model factors, and mention Fig1 again.}] The amalgamation of the prior distribution and the statistical model can be referred to as the generative model.[reference] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
+The amalgamation of the prior distribution and the statistical model can be referred to as the generative model.[reference] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
 
@@
 & = p(\mu)p(\sigma^2|\mu)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{(A,t)})p(n_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) \\
-& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(O_t|\lambda_t, N)p(p_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) 
+& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(O_t|\lambda_t, N)p(p_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}),
  \end{aligned}\right.
 \end{equation}
-
 where $\theta = (\mu, ,\sigma^2,\lambda_t, O_t, p_{(A,t)})$ and $ Y = ( N, n_{(A,t)}, x_{(A,t)}, U_t)$.
@@
 Simulation can be used to predict the performance of the Bayesian models.
-In order to measure the performance of the simplified model, a data set was simulated and the process is described in this chapter.
+In order to measure the performance of the simplified model, a data set was simulated and the process is described in this section.
 
@@
 In terms of the survey data, it is assumed that  we have 12 survey data sets where each survey data set was collected for each month .
-The true value of $p_{A,t}$ was set as $p_{A,t}=0.8$ and the survey sample size $n_{A,t}$ was set as $n_{A,t}=1000$ for every months (t=1:T).
+The true value of $p_{A,t}$ was set as $p_{A,t}=0.8$ and the survey sample size $n_{A,t}$ was set as $n_{A,t}=1000$ for every month (t=1:T).
 The $x_{A,t}$ values for each month were independentally generated from the Binomial distribution (\ref{ambulance}).
-In terms of the overdose data, it is assumed that the total number of population were $N = 10000$ the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
+In terms of the overdose data, it is assumed that the total population size was $N = 10000$, and the true values of parameters for the overdose model were $\mu=\log0.05,\ \sigma=1$.
 The vector of $O_t$ was generated following the overdose model (\ref{overdose}).
 The vector of $U_t$ was generated from the Binomial relation of the two variables ($\ref{over_amb}$).
-The two generated vectors have the same length with the survey data (T=12).
-Note that only $U_t$ and $x_{A,t}$ are the only known as the data set in real world and $p_{A,t}$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
+The two generated vectors have the same length as the survey data (T=12).
+Note that only $U_t$ and $x_{A,t}$ are observable in the real-world application and $p_{A,t}$ needs to be estimated first in order to estimate $O_t$ which is the ultimate interest of the research.\\
 
@@
 [Why do we need MCMC? Because we don't have platinum, golden, and silver grails.]\\
-Posterior distribution of the unseen variables, parameters ($\mu, \sigma^2, p_{A,1:T}$) and the latent variables ($\lambda_{1:T}, O_{1:T}$), given the seen variables, $Y$=($N, n_{(A,1:T)} x_{(A,1:T)}, U_{1:T}$), is the key to make any inference for any of the unseen variables, $\theta$.
+The posterior distribution of the unseen variables, parameters ($\mu, \sigma^2, p_{A,1:T}$) and the latent variables ($\lambda_{1:T}, O_{1:T}$), given the seen variables, $Y$=($N, n_{(A,1:T)} x_{(A,1:T)}, U_{1:T}$), is the key to make any inference for any of the unseen variables, $\theta$.
 For the inference about the variable of interest, $O_{1:T}$, the marginal posterior distribution, $p(O_{1:T}|Y)$, is essential which can be obtained by integrating the joint posterior distribution, $p(\theta|Y)$, over all the other unseen variables except $O_{1:T}$.
-However, in general it is severely complicated and almost impossible to perform such multiple integrations because $p(\theta|Y)$ is a complex distribution of high dimensions.
-Moreover, $p(O_{1:T}|Y)$ may not have closed form nor follow a known parametric distribution, which lead the general inference procedure laborious.\\
+However, in general it is severely complicated and almost impossible to perform such multiple integrations because $p(\theta|Y)$ is a complex distribution of high dimension.
+Moreover, $p(O_{1:T}|Y)$ may not have closed-form nor follow a known parametric distribution, which makes the general inference procedure laborious.\\
 Markov Chain Monte Carlo (MCMC) is a detour to obtain posterior samples from $p(\theta|Y)$ without explicitly knowing the distribution.
-MCMC technique allows us to generate Monte Carlo realizations of samples, $\theta_{(i)}$, which are dependent, with the distribution of each sample, $p(\theta_{(i)})$, converging to $p(\theta|Y)$.\cite{paul}.\\
-[general explanation of what is MC and MC, how it works.]
+MCMC technique allows us to generate Monte Carlo realizations of samples, $\theta_{(i)}$, which are dependent, with the distribution of each sample, $p(\theta_{(i)})$, converging to $p(\theta|Y)$.\cite{paul}.
+[general explanation of what is MC and MC, how it works.]\\
 
  Generative model auto pilot conceptunally there's only\\
- one distribution of unseen stuff given the seen stuff we can use the distribution to make whatever inference claims that we need to make, such as point estimation, credibal interval, inference about parameters or latent variables.
+ one distribution of unseen stuff given the seen stuff we can use the distribution to make whatever inference claims that we need to make, such as point estimation, credibal interval, inference about parameters or latent variables.\\
+ Assessing the properties of a target distribution by generating representative random values is a case of a Monte Carlo simulation.
+Any simulation that samples a lot of random values from a distribution is called a Monte Carlo simulation, named after the dice and spinners and shufflings of the famous casino locale.
+This appellation is attributed to the mathematician von Neumann, or sometimes to Metropolis \cite{gill}.\\
+ The Metropolis algorithm is a specific type of Monte Carlo process.
+It generates a random walk such that each step in the walk is completely independent of the steps before the current position.
+The proposed next step has no dependence on where the walk has been before, and the decision to reject or accept the proposed step has no dependence on where the walk has been before.
+Any such process in which each step has no memory for states before the current one is called a (first order) Markov process, and a succession of such steps is a Markov chain.
+The Metropolis algorithm is an example of a Markov chain Monte Carlo (MCMC) process.
+\cite{doing_bayes}\\
 
@@
 
-[\textbf{combine the two paragraph into one.}]
-All examples here were performed in Python 3.7 using the library pyMC (reference) and JAGS (reference).
-Training was performed using No U-Turn Sampling (NUTS) over two chains with 1000 iterations (is it sample size?).
-Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
-Data processing was carried out using the Pandas and SciPy library [reference].
-Data visualization was performed using the libraries Seaborn and Matplotlib [ref].
+All examples here were performed in Python 3.8 with the library pyMC3 \cite{pymc3} and also in  R 3.6.2 \cite{r_lan} with the JAGS 4.3.0\cite{jags} library.
+Figures in this section Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
 Code for all examples in this study are provided.
-
-The result from the simple case scenario is illustrated below.
-Markov chain Monte Carlo packege PYMC3 3.8\cite{pymc3} was used  with Python 3.7 for the simulation.
-Figures were generated by Python 3.7.
-The results were verified with JAGS 4.3.0 and R 3.6.2.
 
@@
 Figure \ref{pst_ot} is the boxplot of posterior samples of $O_t$.
-It is shown that our posterior estimates of  $O_t$ are fairly accurate since (1) the boxplots contain actual values of $O_t$ within their interquartile range (IQR) and (2) the ranges of IQR and 95\% range seem narrow.
+It is shown that our posterior estimates of  $O_t$ are fairly accurate since (1), the boxplots contain actual values of $O_t$ within their interquartile range (IQR), and (2), the ranges of IQR and 95\% range seem narrow.
 Notice that the range of the boxplot from a higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1, 7, 11)\\
@@
 2000 samples were collected and shown as the histogram.
-It is seen that the plot is bell shaped and not skewed.
+It is seen that the plot is bell-shaped and not skewed.
 The total number of overdoses per year is inferred to be around 10900 to 11000.
@@
 
-\subsection{Initial Result: Contamination of $p_A$ } 
+\subsection{Result: Contamination of $p_A$ } 
 One focus of this research project is to investigate how robust the model is to contaminations of the data sets.
@@
 \begin{aligned}
-o_t \propto u_t \frac{1}{\hat{p}_{A}}
+o_t \propto u_t \frac{1}{\hat{p}_{A}}.
 \end{aligned}
@@
 Figure \ref{contam_ut} gives the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ leads to an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\
+It is seen that none of the contaminations of $p_A$ lead to an contamination effect on $U_t$ but only affect the $O_t$ estimation.\\
 
@@
 The result also matches with our intution.
-Posterior predictive checks use the existing data points twice; firstly it uses the data to obtain the posterior samples of paramaters, and secondly the samples are used to produce posterior predictive samples.
+Posterior predictive checks use the existing data points twice; firstly it uses the data to obtain the posterior samples of parameters, and secondly the samples are used to produce posterior predictive samples.
 Given the fact that there could be an overfitting issue due to using the data twice, the result shown here seems reasonable.
@@
 [????
-I think this should go to the previous of the previous paragraph] This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.
+I think this should go to the previous of the previous paragraph] This is because overdose model (\ref{overdose}) does not affect the result from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.
 

### 2020-04-26 latex/researchpaper.tex
mcmc all done-not sure N ongoing

@@
 Bayesian statistics is a theory in the field of statistics based on the Bayesian interpretation of probability where probability expresses a degree of belief in an event.
-\cite{wiki_bayes} Further explanation is provided at the following section.
-[todo: explain Bayesian statistics a little bit more here]\\
+\cite{wiki_bayes} Further explanation is provided at the following section.\\
 
@@
 Notice that the vector $\theta$ from equation \ref{generic_bayes_thm} includes parameters and latent variables, and hence it contains the total number of overdoses.
+\\
 
@@
  \lambda_{t}^{OD} = \exp(z_{t}),\\
- O_{t} \sim Poi(\lambda_{t}^{OD}N)
+ O_{t} \sim Poi(\lambda_{t}^{OD}N).
  \end{aligned}\right\} 
- \text{(overdose model).} 
+ \text{(overdose model)} 
  \end{equation}
@@
 Variables that need prior distributions are indicated with asterisks.
-The arrows indicate which variables affects which variables; for example, there is an arrow from $(\mu, \sigma^2)$ to $\lambda_{1:T}$ since $\mu, \sigma^2$ are the parameters of $\lambda_{1:T}$, and the two arrows from  $\lambda_{1:T}$ and $N$ to $O_{1:T}$ indicate that both $N$ and $\lambda_{1:T}$ affects the value of $O_{1:T}$ since the product of $\lambda_{1:T}$ and $N$ is the parameter of the distribution of $O_{1:T}$.
+The arrows indicate which variables affects which variables; for example, there is an arrow from $(\mu, \sigma^2)$ to $\lambda_{1:T}$ since $\mu, \sigma^2$ are the parameters of $\lambda_{1:T}$, and the two arrows from  $\lambda_{1:T}$ and $N$ to $O_{1:T}$ indicate that both $N$ and $\lambda_{1:T}$ affects the value of $O_{1:T}$ since the product of $\lambda_{1:T}$ and $N$ is the parameter of the distribution of $O_{1:T}$.\\
 
@@
 \subsection{Generative Models}
-The amalgamation of the prior distribution and the statistical model can be referred to as the generative model.[reference] A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
+The amalgamation of the prior distribution and the statistical model can be referred to as the generative model.\cite{paul} A generative model described in figure \ref{simple_draw} at a month $t$ can be written as follows given some a priori assumptions:
 
@@
 Note that as a function of unknown variables, the function from the second line of the equation \ref{gen_model} is proportional to the function from the third line of the equation since $N$ and $n_{(A,t)}$ are presumed to be known and constants such that the density functions of the known variables are merely constants.
-Therefore the third line of the equation is the ultimate generative model of interest.
+Therefore the third line of the equation is the ultimate generative model of interest.\\
+
+Conceptunally, generative model yields the posterior distribution $p(\theta|Y)$ on auto pilot since there exists only one distribution of $\theta$ given $Y$.
+we can use the distribution to make any inference claims that we need to make about parameters or latent variables such as point estimation and credibal interval.\\
 
@@
 
-[Why do we need MCMC? Because we don't have platinum, golden, and silver grails.]\\
-The posterior distribution of the unseen variables, parameters ($\mu, \sigma^2, p_{A,1:T}$) and the latent variables ($\lambda_{1:T}, O_{1:T}$), given the seen variables, $Y$=($N, n_{(A,1:T)} x_{(A,1:T)}, U_{1:T}$), is the key to make any inference for any of the unseen variables, $\theta$.
-For the inference about the variable of interest, $O_{1:T}$, the marginal posterior distribution, $p(O_{1:T}|Y)$, is essential which can be obtained by integrating the joint posterior distribution, $p(\theta|Y)$, over all the other unseen variables except $O_{1:T}$.
+The posterior distribution of the unseen variables-parameters ($\mu, \sigma^2, p_{A,1:T}$) and the latent variables ($\lambda_{1:T}, O_{1:T}$) given the seen variables, $Y$=($N, n_{(A,1:T)} x_{(A,1:T)}, U_{1:T}$), is the key to make any inference for any of the unseen variables $\theta$.
+For the inference about the variable of interest $O_{1:T}$, the marginal posterior distribution $p(O_{1:T}|Y)$ is essential which can be obtained by integrating the joint posterior distribution $p(\theta|Y)$ over all the other unseen variables except $O_{1:T}$.
 However, in general it is severely complicated and almost impossible to perform such multiple integrations because $p(\theta|Y)$ is a complex distribution of high dimension.
 Moreover, $p(O_{1:T}|Y)$ may not have closed-form nor follow a known parametric distribution, which makes the general inference procedure laborious.\\
+
 Markov Chain Monte Carlo (MCMC) is a detour to obtain posterior samples from $p(\theta|Y)$ without explicitly knowing the distribution.
-MCMC technique allows us to generate Monte Carlo realizations of samples, $\theta_{(i)}$, which are dependent, with the distribution of each sample, $p(\theta_{(i)})$, converging to $p(\theta|Y)$.\cite{paul}.
-[general explanation of what is MC and MC, how it works.]\\
+A Markov chain is a stochastic model describing a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
+In continuous-time, it is known as a Markov process.
+\cite{markov} Monte Carlo refers to the way how the inference of the target distribution is proceeded.
+Assessing the properties of a target distribution by generating representative random values is a case of a Monte Carlo simulation and any simulation that samples a lot of random values from a distribution is called a Monte Carlo simulation.
+\cite{gill}.
+MCMC technique allows us to generate Monte Carlo realizations of samples $\theta_{(i)}$ which are dependent, with the distribution of each sample $p(\theta_{(i)})$ converging to $p(\theta|Y)$.\cite{paul}.
 
- Generative model auto pilot conceptunally there's only\\
- one distribution of unseen stuff given the seen stuff we can use the distribution to make whatever inference claims that we need to make, such as point estimation, credibal interval, inference about parameters or latent variables.\\
- Assessing the properties of a target distribution by generating representative random values is a case of a Monte Carlo simulation.
-Any simulation that samples a lot of random values from a distribution is called a Monte Carlo simulation, named after the dice and spinners and shufflings of the famous casino locale.
-This appellation is attributed to the mathematician von Neumann, or sometimes to Metropolis \cite{gill}.\\
+\paragraph{Algorithms}
+
+Gibbs sampling is a type of Metropolis algorithm that is used in JAGS.
 The Metropolis algorithm is a specific type of Monte Carlo process.
@@
 The proposed next step has no dependence on where the walk has been before, and the decision to reject or accept the proposed step has no dependence on where the walk has been before.
-Any such process in which each step has no memory for states before the current one is called a (first order) Markov process, and a succession of such steps is a Markov chain.
-The Metropolis algorithm is an example of a Markov chain Monte Carlo (MCMC) process.
-\cite{doing_bayes}\\
+The Metropolis algorithm is an example of a MCMC process.
+\cite{doing_bayes} \\
 
-[Two types of bronze grails, NUTS, metropolis]
+The No-U-Turn Sampler (NUTS) is a Hamiltonian Monte Carlo (HMC) Method that is used in pyMC3.
+HMC is a MCMC algorithm that avoids the random walk behavior and sensitivity to correlated parameters that plague many MCMC methods by taking a series of steps informed by first-order gradient information.
+These features allow it to converge to high-dimensional target distributions much more quickly than simpler methods such as random walk Metropolis or Gibbs sampling.
+However HMC’s performance is highly sensitive to two user-specified parameters: a step size and a desired number of steps.
+\cite{hoffman2011nouturn} Moreover, NUTS struggles to samples discrete random variables and hence it is recommended to approximate discrete random variables to continuous ones.
+The Poisson distribution from \ref{overdose} was approximated to Gamma distribution with the same parameter ($\lambda_{t}^{OD}N$) in pyMC3.
 
-[explain all the figures below was derived by NUTS, and python]
+\paragraph{Assessing Convergence}
 
-The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
-The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived.
-\textbf{[Further explanation is going to be added.]} [attempt:brief explaination of MCMC and related topics]
+While independent samples of $\theta$ from the posterior distribution $p(\theta|Y)$ is of interest for any inference purpose, samples of $\theta_{(i)}$ generated by MCMC technique are (1), dependent samples and (2), the distribution of each sample $p(\theta_{(i)})$ are not exactly but only converges to $p(\theta|Y)$.
+It is essential to measure how well Monte Carlo realizations of samples approximate to the possible samples from the independent and identical posterior distribution.
+There are a few steps to examine the level of convergence.\\
+
+First, it is mandatory to throw away some initial samples of Markov Chain realizations.
+Note that a value of $\theta_{(i)}$  from a lot of iteration better represents a sample from the posterior distribuion than some $\theta_{(j)}$ value from some early period of iterations ($i > j$).
+Since the convergence takes enough number of iterations, it is necessary to discard a chunk of the iterations from $\theta_{(1)}$ to $\theta_{(n)}$ for a large number $n$ enough to obtain the proper samples to represent the samples of interest from the posterior distribution.
+The period of iterations to be discarded is called burn-in period.
+\\
+
+Secondly, it is strongly recommended to have multiple starting points to assure the convergence.
+A visual way to check the convergence is to see a plot of  iteration numbers in x axis and their values for each iteration in y axis (trace plot).
+Note that $p(\theta_{(i)})$ converges to  $p(\theta|Y)$ regardless of the initial position of $\theta_{(1)}$, but the location of a initial position affects the number of iterations of MCMC algorithm for the convergence.
+Looking at only a single chain of iterations and their values may mislead us to believe that the number of iterations is enough while the trace of the values are staying at a local area before it reaches to the target distribution.
+By having more than one initial points and checking that the chains get mixed well and are stationary at a certain iteration point, we can verify that the algorithm reaches the desirable target distribution.
+\\
+
+Lastly, numerical measures such as $\hat{R}$ and effective sample size can help us to examine the degree of convergence as well as looking at the trace plot of multiple chains of iterations.
+$\hat{R}$ is a measure of convergence where values for $\hat{R}$ near 1 suggest convergence.
+Effective sample size is a crude measure of the number of independent (effective) sample size from the total number of samples generated from a certain Markov chain.
+Both criteria is accessible after a generative model is fitted from a summary report of MCMC packages.
+The values of $\hat{R}$ and effective sample size in this report are obtainaed from JAGS.
+\\
 
@@
 
-All examples here were performed in Python 3.8 with the library pyMC3 \cite{pymc3} and also in  R 3.6.2 \cite{r_lan} with the JAGS 4.3.0\cite{jags} library.
-Figures in this section Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
+The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
+All examples here were performed in Python 3.8 with the library pyMC3 \cite{pymc3} and also in R 3.6.2 \cite{r_lan} with the JAGS 4.3.0 \cite{jags} library.
+All visualizations were the results from Python and pyMC3.
+Fitting was performed on a GHz Intel Core i5 with 8GM of LPDD3 RAM and typically had wall times under ten minutes.
 Code for all examples in this study are provided.
@@
 	\subfloat[ppc of $p_A$]{\includegraphics[width=.50\columnwidth]{early_r_ppc1_xt}\label{ppc_xt}}
-	\caption[two early result box plots:]{Boxplots of posterior predictive samples of $U_t$ and $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value.
+	\caption[Initial result: box plots of predictive posterior samples of $U_t$ and $x_A$]{Boxplots of posterior predictive samples of $U_t$ and $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value.
 The simulated values are shown as a red dots.}
@@
 	\subfloat[ppc of $O_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-o_t}\label{over_ot}}
-	\caption[two early result box plots:]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $O_t$ values are shown as red dots.}
+	\caption[Initial result: box plot of posterior samples of $O_t$ ]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.
+The actual data point of simulated $O_t$ values are shown as red dots.}
 	\label{contam_ot}
@@
 	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-u_t}\label{over_ut}}
-	\caption[two early result box plots:ut]{Boxplot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
+	\caption[Initial result: boxplot of posterior samples of $U_t$]{Box plot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
 The actual data point of simulated $U_t$ values are shown as red dots.}
@@
 	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
-	\caption[two early result box plots:xa]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.
+	\caption[Initial result: b	oxplot of posterior samples of $x_A$]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.
 The actual data point of simulated $x_A$ values are shown as red dots.}

### 2020-04-27 Rcodes/JAGS_overdoses.Rmd
conclusion draft done, wating for the pymc3

@@
 if (!require(tinytex)) install.packages("tinytex", dependencies = TRUE)
-
+if (!require(MCMCvis)) install.packages("MCMCvis", dependencies = TRUE)
 library('rjags')
@@
 library('tinytex')
-
+library('MCMCvis')
 ```
@@
 # guess it's getting posterior samples ?
+```
+
+```{r}
+MCMCsummary(samples)
+# MCMCtrace(samples)
 ```
@@
 n_a <- 1000
-# N <- 10000
+# N <- 10000 it is to be deleted
 u_t <- df$u_t
@@
 ```{r}
-test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 10000)
-interations = 10000
-burnin= floor(interations/2)
-summary(window(test), start=burnin)
+test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 10000000)
+MCMCsummary(test)
+# interations = 1000000
+# burnin= floor(interations/2)
+# summary(window(test), start=burnin)
 # head(test)
 ```
+
+```{r}
+MCMCtrace(test)
+```
+
 ```{r}
@@
 ```{r}
-plot(N )
-```
-
-```{r}
 ggplot(data = as.data.frame(N), aes(x= N)) +geom_histogram()+ geom_vline(xintercept =10000)

### 2020-04-27 latex/researchpaper.tex
conclusion draft done, wating for the pymc3

@@
 \doublespacing
-
-\begin{abstract}
-
-\end{abstract}
 
@@
 Now the estimated $p_{A,t}$ can help to infer $O_t$ if there is a data regarding the number of  ambulance-attended overdoses.
-Let \(U_t\) be the number of ambulance-attended overdoses at a time point $t$.
+Let \(U_t\) be at a time point $t$.
 In general, the data of ambulance-attended overdoses \(U_t\) can be obtained.
@@
 \label{nonin_prior_amb}
-p(p_A) \sim Beta(1,1)
+p(p_{A,t}) \sim Beta(1,1)
 \text{			noninformative prior for ambulance model}
@@
 Figure \ref{hst_ot} is the histogram of posterior samples of the total number of overdoses $O_+$ per year.
-2000 samples were collected and shown as the histogram.
+3000 samples were collected and shown as the histogram.
 It is seen that the plot is bell-shaped and not skewed.
-The total number of overdoses per year is inferred to be around 10900 to 11000.
-[mention about credible interval] \\
+The 95\% credible interval of the total number of overdoses per year is between 10741 and 11164.\\
 
@@
 This is really a self-consistency check: an observed discrepancy can be due to model misfit or chance.
-\cite{bda_galman}
-[attempted: brief explannation of what is PPC]\\
+\cite{bda_galman}\\
 
@@
 	\subfloat[ppc of $U_t$]{\includegraphics[width=.50\columnwidth]{Figures/early_r_ppc1_ut.png}\label{ppc_ut}}
-	\subfloat[ppc of $p_A$]{\includegraphics[width=.50\columnwidth]{early_r_ppc1_xt}\label{ppc_xt}}
-	\caption[Initial result: box plots of predictive posterior samples of $U_t$ and $x_A$]{Boxplots of posterior predictive samples of $U_t$ and $x_A$ (1000 samples) with the actual data point of simulated $U_t$,$x_A$ value.
+	\subfloat[ppc of $x_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_r_ppc1_xt}\label{ppc_xt}}
+	\caption[Initial result: box plots of predictive posterior samples of $U_t$ and $x_{A,t}$]{Boxplots of posterior predictive samples of $U_t$ and $x_{A,t}$ (1000 samples) with the actual data point of simulated $U_t$,$x_{A,t}$ value.
 The simulated values are shown as a red dots.}
@@
 
-Figure \ref{ppc_xt} is the boxplots of posterior predictive samples of $x_A$.
-It is shown that the posterior predictive estimates of $x_A$ are tolerably accurate since (1), the boxplots contain actual values of $O_t$ within their lines connecting the maximum and the minimum, and (2), the ranges of IQR seem narrow.
+Figure \ref{ppc_xt} is the boxplots of posterior predictive samples of $x_{A,t}$.
+It is shown that the posterior predictive estimates of $x_{A,t}$ are tolerably accurate since (1), the boxplots contain actual values of $O_t$ within their lines connecting the maximum and the minimum, and (2), the ranges of IQR seem narrow.
 \\
 
-\subsection{Result: Contamination of $p_A$ } 
+\subsection{Result: Contamination of $p_{A,t}$ } 
 One focus of this research project is to investigate how robust the model is to contaminations of the data sets.
-The first inspection is to check an impact of a contamination of $p_A$; what would happen if the estimation of $p_A$ is biased?
-It is assumed that the survey data gives us a wrong estimate of $p_A$ such that it would be underestimated or overestimated.
-We then want to see how the biased estimation of $p_A$ affects the estimate of $O_t$, the total number of overdoses.\\
+The first inspection is to check an impact of a contamination of $p_{A,t}$; what would happen if the estimation of $p_{A,t}$ is biased?
+It is assumed that the survey data gives us a wrong estimate of $p_{A,t}$ such that it would be underestimated or overestimated.
+We then want to see how the biased estimation of $p_{A,t}$ affects the estimate of $O_t$, the total number of overdoses.\\
 
 Both underestimation and overestimation were considered for the analysis.
-In terms of underestimation, the simulated survey data ($n_A, x_A$) was generated with $p_A=0.6$ while the true value of $p_A$ is 0.8, and all the other assumptions hold as before.
-That is, $x_A$ is generated from $x_A \sim Bin(n_A, 0.6)$, while $U_t$ is generated from $U_t \sim Bin(O_t, 0.8)$ for every $t$.
-For overestimation, the simulated survey data was generated with $p_A=0.9$ while the true value of $p_A$ is 0.8, and all the other assumptions hold the same.
+In terms of underestimation, the simulated survey data ($n_{A,t}, x_{A,t}$) was generated with $p_{A,t}=0.6$ while the true value of $p_{A,t}$ is 0.8 for all $t$ ($t=1,2,3, ..., 12$), and all the other assumptions hold as before.
+That is, $x_{A,t}$ is generated from $x_{A,t} \sim Bin(n_{A,t}, 0.6)$, while $U_t$ is generated from $U_t \sim Bin(O_t, 0.8)$ for every $t$.
+For overestimation, the simulated survey data was generated with $p_{A,t}=0.9$ while the true value of $p_{A,t}$ is 0.8, and all the other assumptions hold the same.
 \\
@@
 	\centering
-	\subfloat[ppc of $O_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-o_t.png}\label{under_ot}}
-	\subfloat[ppc of $O_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-o_t}\label{over_ot}}
+	\subfloat[ppc of $O_t$: underestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-o_t.png}\label{under_ot}}
+	\subfloat[ppc of $O_t$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-o_t}\label{over_ot}}
 	\caption[Initial result: box plot of posterior samples of $O_t$ ]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.
@@
 
-Figure \ref{contam_ut} gives the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ lead to an contamination effect on $U_t$ but only affect the $O_t$ estimation.\\
+Figure \ref{contam_ut} gives the boxplots of posterior predictive samples where $p_{A,t}$ is underestimated and overestimated respectively from the survey data.
+It is seen that none of the contaminations of $p_{A,t}$ lead to an contamination effect on $U_t$ but only affect the $O_t$ estimation.\\
 
-The possible explanation is that the contaminations of  $p_A$ and $O_t$ (from $p_A$) cancel out the bias component as combined so that $U_t$ has no bias.
+The possible explanation is that the contaminations of  $p_{A,t}$ and $O_t$ (from $p_{A,t}$) cancel out the bias component as combined so that $U_t$ has no bias.
 Notice that $U_t$ is a likelihood (data set) of $O_t$ estimation.
-Hence it plays the role of $y$ in Bayes theorem (\ref{bayes_thm}) and the likelihood, $U_t$, is used to obtain the biased posterior samples of $O_t$ with the contaminated estimates of $p_A$.
-Here the pair of $(O_t, P_A)$ are both biased and the pair is fitted by the likelihood, $U_t$.
+Hence it plays the role of $Y$ in Bayes theorem (\ref{generic_bayes_thm}) and the likelihood, $U_t$, is used to obtain the biased posterior samples of $O_t$ with the contaminated estimates of $p_A$.
+Here the pair of $(O_t, P_{A,t})$ are both biased and the pair is fitted by the likelihood, $U_t$.
 To obtain the posterior predictive samples, MCMC algorithm help obtain posterior samples of biased $O_t$ and then the samples will be combined with biased $p_A$ such that the new samples of $\tilde{U_t}$ from $p(\tilde{U_t}|{U_t})$ is obtained.
@@
 	\centering
-	\subfloat[ppc of $U_t$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-u_t.png}\label{under_ut}}
-	\subfloat[ppc of $U_t$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-u_t}\label{over_ut}}
+	\subfloat[ppc of $U_t$: underestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-u_t.png}\label{under_ut}}
+	\subfloat[ppc of $U_t$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-u_t}\label{over_ut}}
 	\caption[Initial result: boxplot of posterior samples of $U_t$]{Box plot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
@@
 
-Figure \ref{contam_xt}  gives the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data.
-It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots.
+Figure \ref{contam_xt}  gives the boxplots of posterior predictive samples of $x_{A,t}$ where $p_{A,t}$ is underestimated and overestimated respectively from the survey data.
+It is seen that none of the contaminations of $p_{A,t}$ leads an effect of contamination on $x_{A,t}$; the actual simulated points (red dots) are close to the medians from the two boxplots.
 However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900.
-[????
-I think this should go to the previous of the previous paragraph] This is because overdose model (\ref{overdose}) does not affect the result from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.
+This implies that the posterior predictive samples match the biased given data sets (with $p_{A,t}=0.6$ and $p_{A,t}=0.9$ respectively) but the samples cannot estimate the unbiased true value of $p_{A,t}$, which is 0.8.
+This is because overdose model (\ref{overdose}) does not affect the result from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model as it is seen from figure \ref{simple_draw}.
 
@@
 	\centering
-	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-x_a.png}\label{under_xa}}
-	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
-	\caption[Initial result: b	oxplot of posterior samples of $x_A$]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.
-The actual data point of simulated $x_A$ values are shown as red dots.}
+	\subfloat[ppc of $x_{A,t}$: underestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-x_a.png}\label{under_xa}}
+	\subfloat[ppc of $x_{A,t}$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
+	\caption[Initial result: b	oxplot of posterior samples of $x_{A,t}$]{Boxplot of posterior samples of $x_{A,t}$ (1000 samples) where survey data is contaminated.
+The actual data point of simulated $x_{A,t}$ values are shown as red dots.}
 	\label{contam_xt}
@@
 
-\subsection{Variying the Assumptions}
-The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality.
-The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.
+\subsection{Result: Unobservable $N$}
+In practice, it is reasonable to assume that investigators only have a general idea about the total population size of the reigion rather that they know the exact value of the total population size.
+The assumption of the known value of the total population size for a region, $N$, from the previous section changes to be unknown in this section to see the impact of the likelihood over the posterior distributions of variables of interest.
+The generative model at a month $t$ can be written as follows given some a priori assumptions:
 
+\begin{equation}
+\label{gen_model_unk_N}
+\left.\begin{aligned}
+p(\theta, Y) & =p(\mu,\sigma^2,\lambda_t, N, O_t, p_{(A,t)}, n_{(A,t)}, x_{(A,t)}, U_t) \\
+& = p(\mu)p(\sigma^2|\mu)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{(A,t)})p(n_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}) \\
+& \propto p(\mu)p(\sigma^2)p(\lambda_t|\mu, \sigma^2)p(N)p(O_t|\lambda_t, N)p(p_{(A,t)})p(x_{(A,t)}|p_{(A,t)},n_{(A,t)})p(U_t|O_t, p_{(A,t)}),
+\end{aligned}\right.
+\end{equation}
+
+where $\theta = (\mu, ,\sigma^2,\lambda_t, O_t, p_{(A,t)}, N)$ and $ Y = (n_{(A,t)}, x_{(A,t)}, U_t)$.
+Notice that $N$ is an element of $\theta$ instead of an element of $Y$ and the model contains a prior distribution of $N$.
+The chain rule for random variables \cite{chain_rule} is used to the joint distribution of all variables at the first line of equation \ref{gen_model_unk_N} such that the second line of the equation is derived.
+A priori independence assumptions and conditional independence assumptions were presumed so that the second line of the equation is simplified.
+The model is shown graphically in figure \ref{draw_N} where $N$ is a latent variable with a prior distribution needed.\\
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=100mm,scale=0.3]{Figures/draw_N}
+	\caption{Generative model with unoberservable $N$}
+	\label{draw_N}
+\end{figure}
+
+Suppose we have some information about the population size such that we transform it into a prior distribution; the prior distribution of $N$ should reflect our prior knowledge about $N$ which we have even before we encounter the data set.
+As a start,  $N$ is assumed to follow symmetric prior distribution:
+
+\begin{equation}
+\label{in_prior_N}
+N \sim Normal(10000,1000)
+\text{ (informative prior for the total population size),}
+\end{equation}
+
+where the prior information suggests that $N$ is most likely around 10000 and $\pm 20\%$ of the point estimation covers  95\% of the symmetric support of $N$.
+\subsubsection{Posterior Distribution}
+Both MCMC algorithms from pyMC3 and JAGS suffered from relatievely long running time of the algorithms and the convergence issue of the trace of $N$.
+From the trace plot of $N$, two chains of the variable covered a common distribution but neither sequence appeared stationary after 10000000 iterations and the effective sample size for N was 56271.
+Figure \ref{trace_N} shows the trace plot and the posterior density plot of N.
+It is shown that the two chains cross over each other multiple times but they do not seem stationary even with 10000000 iterations.
+The posterior distribution seems to be roughly symmetric at the center 10000.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{Figures/not_mixed_N.png}
+	\caption{Trace plot of $N$}
+	\label{trace_N}
+\end{figure}
+
+The reason of having a mixing issue is because too much inference has been asked to the data set. ~
+
+Posterior samples of $O_t$ was accurate as before.
+
+\subsubsection{Posterior Predictive Check}
+\iffalse
+could vary over time or it can be staritified for a better realization of the reality. The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples. 
+\fi
+
+\section{Conclusion}
+
+Bayesian statistics and generative model is an outstanding approach for statistical inference which incorporate available data sets and our  prior knowlede and information.
+Bayesian model can cope with prior research outcomes as our prior information whereas frequentist approach cannot absorb the knowledge in their models.
+Furthermore, Bayesian modeling is intuitive to us since Bayesian viewpoint to probability corresponds to our conceptual perspective about probability.
+\\
+
+However, Bayesian models have their own weaknesses.
+In cases where the problems are immensely complicated, the models suffer much more than the Frequentist models from long computational time of MCMC algorithms.
+This issue has been being relived as the general computing power of machines has been improved and decent MCMC algorithms have been developed to reduce waiting time although patient waiting is still a virtue to Bayesian statisticians.\\
+
+Note that the robustness of models shoud be confirmed.
+If there exist some contaminations of data sets, the result of posterior estimations may not be trustworthy and therefore it is needed to confirm how robust the result of models is from the contaimiations.
+Likewise, it is generally recommended to apply more than one MCMC algorithms to the generative model to confirm that Monte Carlo realizations of  samples are robust to different types of the algorithms.\\
+
+Lastly, notice that it is not reasonable to provide few data sets and ask Bayesian generative models to yield decent results; the models are not a god-like solver for every qnatitative problems.
+It is shown from the previous section that asking about the total population size $N$ is too much to ask for the model while the survey data $(n_{A,1:T}, x_{A,1:T})$ and the data of the number of ambulance-attended overdoses $(U_{1:T})$ are only provided.	\\
+
+ nevertheless, the bayesian generative model is 
 \bibliographystyle{alpha}

### 2020-04-27 Rcodes/JAGS_overdoses.Rmd
figures lastly added

@@
 MCMCtrace(test) 
-```
-
-```{r}
-plot(test)
+# better than #plot(test)
 ```
@@
 N <- pst_mtx.N[,1]
+head(N)
 length(N)
@@
 ```{r}
-ggplot(data = as.data.frame(N), aes(x= N)) +geom_histogram()+ geom_vline(xintercept =10000)
+
+hist(N)
 ```
+
+## get the $O_t$, $U_t$, $x_{A,t}$ ommitted for now.
 

### 2020-04-27 latex/researchpaper.tex
figures lastly added

@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{Figures/hist_N.png}
+	\caption{Boxplot of posterior samples of $N$ (20000 samples for each month)}
+	\label{hist_unk_n}
+\end{figure}
+
 Posterior samples of $O_t$ was accurate as before.
+[mention that although the mixing was not possible with the maximum capacity of the computer, at least the proper prior seems to help a lot for the estimation of $O_t$ which depends on the N.
+][put the figure and briefly explain that the the estimation for the variable is not hugely affected.]
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{Figures/unknown_N_ot.png}
+	\caption{Boxplot of posterior samples of $O_t$ (20000 samples for each month) with simulated $O_t$ values as red dots.}
+	\label{pst_ot_unk_n}
+\end{figure}
 
 \subsubsection{Posterior Predictive Check}
+
+[put the figures and briefly explain that the the estimation for the variables are not hugely affected.]
+\begin{figure}[htb]
+	\centering
+	\subfloat[ppc of $U_t$]{\includegraphics[width=.50\columnwidth]{Figures/unk_N_ppc1_ut.png}\label{ppc_ut_unk_n}}
+	\subfloat[ppc of $x_{A,t}$]{\includegraphics[width=.50\columnwidth]{unk_N_ppc1_xt}\label{ppc_xt_unk_n}}
+	\caption[Initial result: box plots of predictive posterior samples of $U_t$ and $x_{A,t}$]{Boxplots of posterior predictive samples of $U_t$ and $x_{A,t}$ (1000 samples) with the actual data point of simulated $U_t$,$x_{A,t}$ value.
+The simulated values are shown as a red dots.}
+
+\end{figure}
+
 \iffalse
@@
 
- nevertheless, the bayesian generative model is 
 \bibliographystyle{alpha}

### 2020-04-27 latex/researchpaper.tex
I am dying

@@
 	\includegraphics[width=200mm,scale=0.5]{Figures/simple_drawing}
-	\caption{Example of estimating the total number of overdoses indirectly given two data sets: ambulance attended overdoses and survey data.}
+	\caption[Example of estimating the total number of overdoses]{Example of estimating the total number of overdoses indirectly given two data sets: ambulance attended overdoses and survey data.}
 	\label{simple_draw}
@@
 	\includegraphics[width=1\linewidth]{Figures/earlyresult1_ot.png}
-	\caption{Boxplot of posterior samples of $O_t$ (2000 samples for each month) with simulated $O_t$ values as red dots.}
+	\caption[Initial result: boxplot of posterior samples of $O_t$]{Boxplot of posterior samples of $O_t$ (2000 samples for each month) with simulated $O_t$ values as red dots.}
 	\label{pst_ot}
@@
 	\includegraphics[width=1\linewidth]{Figures/hist_sum_ot.png}
-	\caption{Histogram of posterior samples of total overdoses per year}
+	\caption[Initial result: histogram of posterior samples of total overdoses per year]{Histogram of posterior samples of total overdoses per year}
 	\label{hst_ot}
@@
 	\subfloat[ppc of $O_t$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-o_t}\label{over_ot}}
-	\caption[Initial result: box plot of posterior samples of $O_t$ ]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.
+	\caption[Contaminated $p_{A,t}$: boxplot of posterior samples of $O_t$ ]{Boxplot of posterior samples of $O_t$ (1000 samples) where survey data is contaminated.
 The actual data point of simulated $O_t$ values are shown as red dots.}
@@
 	\subfloat[ppc of $U_t$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-u_t}\label{over_ut}}
-	\caption[Initial result: boxplot of posterior samples of $U_t$]{Box plot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
+	\caption[Contaminated $p_{A,t}$: boxplot of posterior samples of $U_t$]{Box plot of posterior samples of $U_t$ (1000 samples) where survey data is contaminated.
 The actual data point of simulated $U_t$ values are shown as red dots.}
@@
 	\subfloat[ppc of $x_{A,t}$: overestimated $p_{A,t}$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
-	\caption[Initial result: b	oxplot of posterior samples of $x_{A,t}$]{Boxplot of posterior samples of $x_{A,t}$ (1000 samples) where survey data is contaminated.
+	\caption[Contaminated $p_{A,t}$:: boxplot of posterior samples of $x_{A,t}$]{Boxplot of posterior samples of $x_{A,t}$ (1000 samples) where survey data is contaminated.
 The actual data point of simulated $x_{A,t}$ values are shown as red dots.}
@@
 	\includegraphics[width=1\linewidth]{Figures/not_mixed_N.png}
-	\caption{Trace plot of $N$}
+	\caption[Unobserved $N$: trace plot of latent variable $N$]{Trace plot of latent variable $N$}
 	\label{trace_N}
@@
 	\centering
-	\includegraphics[width=1\linewidth]{Figures/hist_N.png}
-	\caption{Boxplot of posterior samples of $N$ (20000 samples for each month)}
+	\includegraphics[width=0.7\linewidth]{Figures/hist_N.png}
+	\caption[Unobserved $N$: boxplot of posterior samples of $N$]{Boxplot of posterior samples of $N$ (20000 samples for each month)}
 	\label{hist_unk_n}
@@
 	\includegraphics[width=1\linewidth]{Figures/unknown_N_ot.png}
-	\caption{Boxplot of posterior samples of $O_t$ (20000 samples for each month) with simulated $O_t$ values as red dots.}
+
+	\caption[Unobserved $N$: boxplot of posterior samples of $O_t$]{Boxplot of posterior samples of $O_t$ (20000 samples for each month) with simulated $O_t$ values as red dots.}
 	\label{pst_ot_unk_n}
@@
 	\subfloat[ppc of $x_{A,t}$]{\includegraphics[width=.50\columnwidth]{unk_N_ppc1_xt}\label{ppc_xt_unk_n}}
-	\caption[Initial result: box plots of predictive posterior samples of $U_t$ and $x_{A,t}$]{Boxplots of posterior predictive samples of $U_t$ and $x_{A,t}$ (1000 samples) with the actual data point of simulated $U_t$,$x_{A,t}$ value.
+	\caption[Unobserved $N$: box plots of predictive posterior samples of $U_t$ and $x_{A,t}$]{Boxplots of posterior predictive samples of $U_t$ and $x_{A,t}$ (1000 samples each) with the actual data point of simulated $U_t$, $x_{A,t}$ value.
 The simulated values are shown as a red dots.}

### 2020-04-27 latex/researchpaper.tex
drafty draft done...

@@
 Both MCMC algorithms from pyMC3 and JAGS suffered from relatievely long running time of the algorithms and the convergence issue of the trace of $N$.
-From the trace plot of $N$, two chains of the variable covered a common distribution but neither sequence appeared stationary after 10000000 iterations and the effective sample size for N was 56271.
-Figure \ref{trace_N} shows the trace plot and the posterior density plot of N.
-It is shown that the two chains cross over each other multiple times but they do not seem stationary even with 10000000 iterations.
+The reason of having the stationarity issue is that too much inference has been asked to the data sets.
+Only two data sets, $U_{1:T}$ and $(n_{A,1:T}, x_{A,1:T})$, are given and the task was to generate posterior samples of 5 random variables, $N, \mu, \sigma^2, \lambda_{1:T}, O_{1:T}, p_{A,1:T}$.
+Figure \ref{trace_N} shows the trace plot and the posterior density plot of $N$.
+From the trace plot of $N$ it is seen that the chains are mixed well but not stationary and thus not convenging; two chains of the variable covered a common distribution but neither sequence appeared stable even after 10000000 iterations.
+The effective sample size for N was only 56271.
 The posterior distribution seems to be roughly symmetric at the center 10000.
-\begin{figure}[h]
+\begin{figure}[htb]
 	\centering
@@
 
-The reason of having a mixing issue is because too much inference has been asked to the data set. ~
+Figure \ref{hist_unk_n} shows the posterior samples of $N$.
+Due to the computational cost and the restriction of computing power of the research device, only 20000 samples were generated from the model.
+The plot reveals that the posterior distribution is centered around the true value, 10000, and symmetric around the center.
+The result is likely to be due to the prior distribution.
+The prior distribution was providing exact and accurate information about $N$ such that the posterior distribution remains the similar shape although there is a convergence issue.
+Further research is recommended to investigate the effect of prior distribution to the posteior distribution of $N$.
 
-\begin{figure}[h]
+\begin{figure}[htb]
 	\centering
@@
 
-Posterior samples of $O_t$ was accurate as before.
-[mention that although the mixing was not possible with the maximum capacity of the computer, at least the proper prior seems to help a lot for the estimation of $O_t$ which depends on the N.
-][put the figure and briefly explain that the the estimation for the variable is not hugely affected.]
-\begin{figure}[h]
+Figure \ref{pst_ot_unk_n} shows that the inference of $O_{1:T}$ as accurate as before.
+This was possible as the proper prior distribution of $N$ yields acceptable estimation of $N$ which dominates the estimation of $O_t$ with $\mu, \sigma^2$ from figure \ref{draw_N}.
+Any distortion of the prior distribution of $N$ may affect the inference result of $O_{1:T}$.
+
+\begin{figure}[htb]
 	\centering
@@
 
-[put the figures and briefly explain that the the estimation for the variables are not hugely affected.]
+Figure \ref{ppc_unk_n} is the boxplot of posterior predictive samples of $U_t$ and $x_{A,t}$.
+It is shown that the posterior predictive estimates of $U_t$ and $x_{A,t}$ are failry accurate for the same reason regarding the accuracy of the posterior distribution of $O_t$.
+
 \begin{figure}[htb]
@@
 The simulated values are shown as a red dots.}
-
+	\label{ppc_unk_n}
 \end{figure}
-
-\iffalse
-could vary over time or it can be staritified for a better realization of the reality. The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples. 
-\fi
 

### 2020-05-01 latex/researchpaper.tex
finally done, my first science research project

@@
 \title{Bayesian Evidence Synthesis:Opioid Crisis}
-\author{Hyeongcheol Park*}
+\author{Hyeongcheol Park}
 
@@
 Bayesian statistics is a theory in the field of statistics based on the Bayesian interpretation of probability where probability expresses a degree of belief in an event.
-\cite{wiki_bayes} Further explanation is provided at the following section.\\
+\cite{wiki_bayes} Further explanation is provided in the following section.\\
 
@@
 To achieve the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation.
-The posterior belief is coming from both of these sources of the information.
+The posterior belief is coming from both of these sources of information.
 Let $\theta$ denote all the unknown variables and let $Y$ denote all the related data sets.
-The following equation express the idea of Bayesian generative models in a mathmatical form:
+The following equation expresses the idea of Bayesian generative models in a mathmatical form:
 \begin{equation}
@@
 \end{equation}
-
 where $p(\cdot)$ is probability density (or mass) function.
@@
 \fi
-In Bayesian concepts, the variable of interest $O_{1:T}$ is a vector of fixed numbers but the numbers are yielded from a certain distribution where the distribution describes knowledge about $O_{1:T}$; that is, each $O_t$ is considered to be a sample of a random variable.
-The simplest conceptual model is to take an underlying log-rate $z_t$ for each month $t$ where $z_t$ is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
+Applying Bayesian concepts, the variable of interest $O_{1:T}$ is a vector of fixed numbers but the numbers are yielded from a certain distribution where the distribution describes knowledge about $O_{1:T}$; that is, each $O_t$ is considered to be a sample of a random variable.
+The simplest conceptual model is to take an underlying log-rate $z_t$ for each month $t$, where $z_t$ is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$.
 \cite{Irvine:modelling} Then let $\lambda_{t}^{OD}$ denote the rate of overdose at month $t$.
-It is assumed that the number of overdoses in the month, $O_t$, follows Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the size of the population in the region of interest is $N$.
+It is assumed that the number of overdoses in the month, $O_t$, follows a Poisson distribution with parameter $\lambda_{t}^{OD}N$ where the size of the population in the region of interest is $N$.
 Note that the population size, N, is unknown and vary over time in reality but it is assumed known and constant for the investigator here for simplicity.
@@
 
-Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ is not known, we assume there was a survey conducted to help to support the inference about  $O_t$.
+Since estimation of $O_t$ is not straightforward as some of the variables ($\mu$, $\sigma$) determining $O_t$ are not known, we assume there was a survey conducted to help to support the inference about  $O_t$.
 The survey asked witnesses of opioid overdoses whether someone called an ambulance, for each month.
@@
 
-Now the estimated $p_{A,t}$ can help to infer $O_t$ if there is a data regarding the number of  ambulance-attended overdoses.
+Now the estimated $p_{A,t}$ can help to infer $O_t$ if there is data regarding the number of  ambulance-attended overdoses.
 Let \(U_t\) be at a time point $t$.
@@
 Figure \ref{simple_draw} shows an example of how the estimation can proceed.
-The variables that can be observed (N, $U_{1:T}$, and $x_{A,1:T}$) are drawn in squares and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, p_{A,1:T} $) that are not observarble from possible data sets are drawn in circles.
+The variables that can be observed (N, $U_{1:T}$, and $x_{A,1:T}$) are drawn in squares and the parameters ($\mu, \sigma^2$) and the latent variables ($\lambda_{1:T}, O_{1:T}, p_{A,1:T} $) that are not observable from possible data sets are drawn in circles.
 Variables that need prior distributions are indicated with asterisks.
@@
 A priori independence assumptions and conditional independence assumptions were presumed so that the second line of the equation is simplified.
-For example, it is asumed that  $\sigma^2$ and $\mu$ are independent to each other and $p_{A,t}$ and also all the variables from overdose model \ref{overdose}, ($\mu, \sigma^2, \lambda_t, N, O_t$), are independent to each other.
+For example, it is asumed that  $\sigma^2$ and $\mu$ are independent of each other and $p_{A,t}$ and also all the variables from overdose model \ref{overdose}, ($\mu, \sigma^2, \lambda_t, N, O_t$), are independent of each other.
 Note that as a function of unknown variables, the function from the second line of the equation \ref{gen_model} is proportional to the function from the third line of the equation since $N$ and $n_{(A,t)}$ are presumed to be known and constants such that the density functions of the known variables are merely constants.
@@
 
-Conceptunally, generative model yields the posterior distribution $p(\theta|Y)$ on auto pilot since there exists only one distribution of $\theta$ given $Y$.
+The generative model yields the posterior distribution $p(\theta|Y)$ on auto pilot, since there exists only one distribution of $\theta$ given $Y$.
 we can use the distribution to make any inference claims that we need to make about parameters or latent variables such as point estimation and credibal interval.\\
@@
 In continuous-time, it is known as a Markov process.
-\cite{markov} Monte Carlo refers to the way how the inference of the target distribution is proceeded.
+\cite{markov} Monte Carlo refers to how inference about the target distribution proceeds.
 Assessing the properties of a target distribution by generating representative random values is a case of a Monte Carlo simulation and any simulation that samples a lot of random values from a distribution is called a Monte Carlo simulation.
@@
 However HMC’s performance is highly sensitive to two user-specified parameters: a step size and a desired number of steps.
-\cite{hoffman2011nouturn} Moreover, NUTS struggles to samples discrete random variables and hence it is recommended to approximate discrete random variables to continuous ones.
+\cite{hoffman2011nouturn} Moreover, NUTS struggles to sample discrete random variables and hence it is recommended to approximate discrete random variables with continuous ones.
 The Poisson distribution from \ref{overdose} was approximated to Gamma distribution with the same parameter ($\lambda_{t}^{OD}N$) in pyMC3.
@@
 First, it is mandatory to throw away some initial samples of Markov Chain realizations.
-Note that a value of $\theta_{(i)}$  from a lot of iteration better represents a sample from the posterior distribuion than some $\theta_{(j)}$ value from some early period of iterations ($i > j$).
-Since the convergence takes enough number of iterations, it is necessary to discard a chunk of the iterations from $\theta_{(1)}$ to $\theta_{(n)}$ for a large number $n$ enough to obtain the proper samples to represent the samples of interest from the posterior distribution.
-The period of iterations to be discarded is called burn-in period.
+Note that a value of $\theta_{(i)}$  from a lot of iteration better represents a sample from the posterior distribution than some $\theta_{(j)}$ value from some early period of iterations ($i > j$).
+Since the convergence takes some number of iterations, it is necessary to discard a chunk of the iterations from $\theta_{(1)}$ to $\theta_{(n)}$ for a large number $n$ enough to obtain the proper samples to represent the samples of interest from the posterior distribution.
+The period of iterations to be discarded is called the burn-in period.
 \\
@@
 Note that $p(\theta_{(i)})$ converges to  $p(\theta|Y)$ regardless of the initial position of $\theta_{(1)}$, but the location of a initial position affects the number of iterations of MCMC algorithm for the convergence.
-Looking at only a single chain of iterations and their values may mislead us to believe that the number of iterations is enough while the trace of the values are staying at a local area before it reaches to the target distribution.
-By having more than one initial points and checking that the chains get mixed well and are stationary at a certain iteration point, we can verify that the algorithm reaches the desirable target distribution.
+Looking at only a single chain of iterations and their values may mislead us to believe that the number of iterations is enough, with the trace of the values are staying at a local area before it reaches to the target distribution.
+By having more than one initial point and checking that the chains mix well and are stationary at a certain iteration point, we can verify that the algorithm reaches the desirable target distribution.
 \\
@@
 Effective sample size is a crude measure of the number of independent (effective) sample size from the total number of samples generated from a certain Markov chain.
-Both criteria is accessible after a generative model is fitted from a summary report of MCMC packages.
-The values of $\hat{R}$ and effective sample size in this report are obtainaed from JAGS.
+Both criteria are accessible after a generative model is fitted from a summary report of MCMC packages.
+The values of $\hat{R}$ and effective sample size in this report are obtained from JAGS.
 \\
@@
 
-The achievement of the simulation was done by Markov chain Monte Carlo (MCMC).
+The implementation of the simulation was done by Markov chain Monte Carlo (MCMC).
 All examples here were performed in Python 3.8 with the library pyMC3 \cite{pymc3} and also in R 3.6.2 \cite{r_lan} with the JAGS 4.3.0 \cite{jags} library.
@@
 Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$.
-It is shown that the posterior predictive estimates of $U_t$ are failry accurate, for the same two reasons regarding the accuracy of the posterior distribution of $O_t$.
+It is shown that the posterior predictive estimates of $U_t$ are accurate, for the same two reasons regarding the accuracy of the posterior distribution of $O_t$.
 It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11).
@@
 This implies that the posterior predictive samples match the biased given data sets (with $p_{A,t}=0.6$ and $p_{A,t}=0.9$ respectively) but the samples cannot estimate the unbiased true value of $p_{A,t}$, which is 0.8.
-This is because overdose model (\ref{overdose}) does not affect the result from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model as it is seen from figure \ref{simple_draw}.
+This is because overdose model (\ref{overdose}) does not affect the result from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model as is seen from figure \ref{simple_draw}.
 
@@
 \subsection{Result: Unobservable $N$}
-In practice, it is reasonable to assume that investigators only have a general idea about the total population size of the reigion rather that they know the exact value of the total population size.
+In practice, it is reasonable to assume that investigators only have a general idea about the total population size of the region rather that they know the exact value of the total population size.
 The assumption of the known value of the total population size for a region, $N$, from the previous section changes to be unknown in this section to see the impact of the likelihood over the posterior distributions of variables of interest.
@@
 \end{equation}
-
 where $\theta = (\mu, ,\sigma^2,\lambda_t, O_t, p_{(A,t)}, N)$ and $ Y = (n_{(A,t)}, x_{(A,t)}, U_t)$.
@@
 \end{equation} 
-
 where the prior information suggests that $N$ is most likely around 10000 and $\pm 20\%$ of the point estimation covers  95\% of the symmetric support of $N$.
 \subsubsection{Posterior Distribution}
-Both MCMC algorithms from pyMC3 and JAGS suffered from relatievely long running time of the algorithms and the convergence issue of the trace of $N$.
-The reason of having the stationarity issue is that too much inference has been asked to the data sets.
+Both MCMC algorithms from pyMC3 and JAGS suffered from relatively long running time of the algorithms and a convergence issue for the trace of $N$.
+The reason of having the stationarity issue is that too much inference has been asked of the data sets.
 Only two data sets, $U_{1:T}$ and $(n_{A,1:T}, x_{A,1:T})$, are given and the task was to generate posterior samples of 5 random variables, $N, \mu, \sigma^2, \lambda_{1:T}, O_{1:T}, p_{A,1:T}$.
@@
 Figure \ref{ppc_unk_n} is the boxplot of posterior predictive samples of $U_t$ and $x_{A,t}$.
-It is shown that the posterior predictive estimates of $U_t$ and $x_{A,t}$ are failry accurate for the same reason regarding the accuracy of the posterior distribution of $O_t$.
+It is shown that the posterior predictive estimates of $U_t$ and $x_{A,t}$ are accurate for the same reason regarding the accuracy of the posterior distribution of $O_t$.
 
@@
 
-Bayesian statistics and generative model is an outstanding approach for statistical inference which incorporate available data sets and our  prior knowlede and information.
-Bayesian model can cope with prior research outcomes as our prior information whereas frequentist approach cannot absorb the knowledge in their models.
-Furthermore, Bayesian modeling is intuitive to us since Bayesian viewpoint to probability corresponds to our conceptual perspective about probability.
+Bayesian statistics and generative model is an outstanding approach for statistical inference which incorporate available data sets and our  prior knowledge and information.
+Bayesian model can cope with prior research outcomes as our prior information whereas frequentist approach cannot absorb such knowledge in their models.
+Furthermore, Bayesian modeling is intuitive to us since the Bayesian viewpoint to probability corresponds to our conceptual perspective about probability.
 \\
@@
 In cases where the problems are immensely complicated, the models suffer much more than the Frequentist models from long computational time of MCMC algorithms.
-This issue has been being relived as the general computing power of machines has been improved and decent MCMC algorithms have been developed to reduce waiting time although patient waiting is still a virtue to Bayesian statisticians.\\
+This issue has been being relived as the general computing power of machines has been improved and decent MCMC algorithms have been developed to reduce waiting time, although patient waiting is still a virtue to Bayesian statisticians.\\
 
 Note that the robustness of models shoud be confirmed.
-If there exist some contaminations of data sets, the result of posterior estimations may not be trustworthy and therefore it is needed to confirm how robust the result of models is from the contaimiations.
-Likewise, it is generally recommended to apply more than one MCMC algorithms to the generative model to confirm that Monte Carlo realizations of  samples are robust to different types of the algorithms.\\
+If there exist some contaminations of data sets, the result of posterior estimations may not be trustworthy and therefore it is necessary to confirm how robust the results of models are to the contaimiations.
+Likewise, it is generally recommended to apply more than one MCMC algorithm to the generative model to confirm that Monte Carlo realizations of  samples are robust to different types of the algorithms.\\
 
-Lastly, notice that it is not reasonable to provide few data sets and ask Bayesian generative models to yield decent results; the models are not a god-like solver for every qnatitative problems.
-It is shown from the previous section that asking about the total population size $N$ is too much to ask for the model while the survey data $(n_{A,1:T}, x_{A,1:T})$ and the data of the number of ambulance-attended overdoses $(U_{1:T})$ are only provided.	\\
+Lastly, notice that it is not reasonable to provide few data sets and ask Bayesian generative models to yield decent results; the models are not a god-like solver for every quantitative problem.
+It is shown in the previous section that asking about the total population size $N$ is too much to ask for the model when only the survey data $(n_{A,1:T}, x_{A,1:T})$ and the data of the number of ambulance-attended overdoses $(U_{1:T})$ are provided.	\\
 
